### 2015-06-07 doc/article.tex
added draft stuff

% Template for PLoS
% Version 3.1 February 2015
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig." instead of "Figure".
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{Multiple RNA structure consensus assignments}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Gabriel C-Parent, % \textsuperscript{1},
Francois Major%\textsuperscript{1},
%with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
%\bf{1} 
Institute for Research in Immunology and Cancer, and Department of Computer Science, Université de Montréal, Montreal, QC, H3C 3J7, Canada
%\\
%\bf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
%\\
%\bf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.
%
%% Additional Equal Contribution Note
%% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.
%
%% Current address notes
%\textcurrency a Insert current address of first author with an address update
%% \textcurrency b Insert current address of second author with an address update
%% \textcurrency c Insert current address of third author with an address update
%
%% Deceased author note
%\dag Deceased
%
%% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.
%
%% Use the asterisk to denote corresponding authorship and provide email address in note below.
* gabriel.c-parent@umontreal.ca

\end{flushleft}
% Please keep the abstract below 300 words

% example of author summary vs abstract forom Week's latest paper
%Abstract
%
%Discovery and characterization of functional RNA structures remains challenging due to deficiencies in de novo secondary structure modeling. Here we describe a dynamic programming approach for model-free sequence comparison that incorporates high-throughput chemical probing data. Based on SHAPE probing data alone, ribosomal RNAs (rRNAs) from three diverse organisms – the eubacteria E. coli and C. difficile and the archeon H. volcanii – could be aligned with accuracies comparable to alignments based on actual sequence identity. When both base sequence identity and chemical probing reactivities were considered together, accuracies improved further. Derived sequence alignments and chemical probing data from protein-free RNAs were then used as pseudo-free energy constraints to model consensus secondary structures for the 16S and 23S rRNAs. There are critical differences between these experimentally-informed models and currently accepted models, including in the functionally important neck and decoding regions of the 16S rRNA. We infer that the 16S rRNA has evolved to undergo large-scale changes in base pairing as part of ribosome function. As high-quality RNA probing data become widely available, structurally-informed sequence alignment will become broadly useful for de novo motif and function discovery.
%
%Author Summary
%
%Despite the clear functional importance of structure in RNA molecules, it remains very difficult to correctly identify and annotate similar RNA structures because their sequences are often poorly conserved even for RNAs that form very similar higher-order structures. A solution is to use a metric that identifies structural motifs but this, too, is difficult because RNA structure modeling based on sequence alone is generally not very accurate. In this work, we use SHAPE chemical probing to obtain model-free information about RNA structure and then exploit this information to align RNAs by sequence. We show for ribosomal RNAs that sequence alignments based on SHAPE experimental information alone are as accurate as those that actually use sequence information. In addition, we identify regions in the 16S ribosomal RNA that form conserved secondary structures that are different from currently accepted models. These differences, rather than being errors, reveal sites of conformational flexibility that may underlie mechanistic functions in ribosome assembly and translation regulation. We anticipate that structure-informed sequence alignment and structure modeling will become broadly useful tools in RNA function analysis.

\section*{Abstract}
RNA secondary structural assignment is an interesting problem. Bla bla bla.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.

\section*{Author Summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

\section*{Introduction}
Lorem ipsum dolor %sit~\cite{bib1}%%
amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq.~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\begin{equation}\label{eq:schemeP} 
D_{coll} = \frac{D_f+\frac{[S]^2}{K_D S_T} D_S} {1+\frac{[S]^2}{K_D S_T}}, 
D_{sm} = \frac{D_f+ \frac{[S]}{K_D} D_S}{1+\frac{[S]}{K_D}},
\end{equation}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis"
\newpage
\section*{Materials and Methods}
\subsection*{Theoretical foundation}

\subsection*{Metaheuristics}

\subsection*{MCCONS}

\paragraph{Metrics}

\subparagraph{Tree Indel distance on base pair tree}

RNA secondary structure is represented as a tree, where nodes are base pairs and unpaired positions are ignored.

\subparagraph{String Edit Distance}
%\subsection*{}

\newpage

% For figure citations, please use "Fig." instead of "Figure".
Nulla mi mi, Fig.~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{figure}[h]
\caption{{\bf Figure Title first bold sentence Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}
Figure Caption Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

\begin{enumerate}
\item{react}
\item{diffuse free particles}
\item{increment time by dt and go to 1}
\end{enumerate}

% Results and Discussion can be combined.
\section*{Results}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat.

\subsection*{Sed ac quam id nisi malesuada congue.}

Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}. 
\subsection*{IRES}

Although IRES tend to fold in very similar shapes,

\subsection*{A more advanced level: tRNA alternative folding}
\paragraph{3rd Level Heading.} Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\section*{Discussion}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}

CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Text}.

\section*{Supporting Information}

% Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
\subsection*{S1 Video}
\label{S1_Video}
{\bf Bold the first sentence.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S1 Text}
\label{S1_Text}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S1 Fig}
\label{S1_Fig}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S2 Fig}
\label{S2_Fig}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S1 Table}
\label{S1_Table}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

%\section*{References}
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here.
% 
%\begin{thebibliography}{10}
%\bibitem{bib1}
%Devaraju P, Gulati R, Antony PT, Mithun CB, Negi VS. Susceptibility to SLE in South Indian Tamils may be influenced by genetic selection pressure on TLR2 and TLR9 genes. Mol Immunol. 2014 Nov 22. pii: S0161-5890(14)00313-7. doi: 10.1016/j.molimm.2014.11.005
%
%\bibitem{bib2}
%Huynen MMTE, Martens P, Hilderlink HBM. The health impacts of globalisation: a conceptual framework. Global Health. 2005;1: 14. Available: http://www.globalizationandhealth.com/content/1/1/14.
%
%\end{thebibliography}

\end{document}

### 2015-06-07 include/ConsensusProblem.h
did minor reformating and removed maximum string length from string edit distance

@@ -1,10 +1,11 @@
 #ifndef CONSENSUSPROBLEM_H
 #define CONSENSUSPROBLEM_H
 
-#include <vector>
-#include <string>
-#include <limits> // for infinity
+#include <limits>
 #include <map>
+#include <string>
+#include <vector>
+
 #include "Tree.h"
 
 struct Range

### 2015-06-07 include/Node.h

@@ -6,19 +6,20 @@
 class Node
 {
 public:
-    // constructors
+    // constructor
     Node(Node* parent=NULL, int label=-1);
 
     // destructor
     ~Node();
 
     // getters and setters
-    void add_child(Node* other);
-    const std::vector<Node*> & get_children();
     Node* get_parent();
     int get_label();
+    const std::vector<Node*> & get_children();
+
     void set_parent(Node* parent);
     void set_label(int label);
+    void add_child(Node* other);
 
     // helper
     int get_leftmost_descendent_label();
@@ -29,8 +30,6 @@ private:
     std::vector<Node*> children;
 };
 
-
-// some useful functions
 
 // node <-> bracket interconversion
 void node_to_bracket(Node* node, std::vector<char>& brackets);

### 2015-06-07 include/ProgressBar.h

@@ -15,8 +15,10 @@ public:
 
     // updates the progress displayed
     void update(float percentage);
+
     // displays 100% status
     void clean();
+
 private:
     size_t _bar_length;
     std::string _process_name;

### 2015-06-07 include/RNA.h

@@ -1,19 +1,19 @@
 #ifndef SRC_CPP_RNA_H_
 #define SRC_CPP_RNA_H_
 
+#include <algorithm>
+
 #include "Tree.h"
 #include "RngStream.h"
 
-#include <algorithm>
 
 // rna secondary structure general helpers (conversion and distances)
-
 
 bool is_valid_dot_bracket(std::string dot_bracket);
 
 std::string only_paired(std::string dot_bracket);
 
-int levenshtein(std::string  a, std::string  b);
+int levenshtein(std::string a, std::string b);
 
 Node* dot_bracket_to_node(std::string  dot_bracket);  // don't forget to free it
 

### 2015-06-07 include/Solution.h

@@ -14,10 +14,11 @@ public:
     Solution(std::vector<int> chosen_genes);
     Solution(std::vector<int> chosen_genes, double score);
     Solution(const Solution& other); // copy constructor
+
     ~Solution();
 
     // getters and setters
-    std::vector<int> get_genes() const;
+    const std::vector<int> & get_genes() const;
     void set_gene(int position, int new_gene);
     void add_gene(int new_gene);
 

### 2015-06-07 include/SolverGA.h

@@ -1,9 +1,8 @@
-#ifndef SOLVERGA_H
-#define SOLVERGA_H
+#ifndef SOLVERGA_H_
+#define SOLVERGA_H_
 
 #include <algorithm>
 #include <vector>
-
 
 #include <assert.h>
 #include <math.h>
@@ -18,7 +17,8 @@
 class SolverGA {
 public:
     // constructors and destructors
-    SolverGA(// data
+    SolverGA(
+             // data
              std::vector< std::vector<double> > distance_matrix,
              std::vector<Range> ranges,
 
@@ -71,5 +71,4 @@ void steepest_improvement(Solution& sol,
                           std::vector<Range> ranges,
                           int max_iter=pow(2, 20));
 
-
-#endif // SOLVERGA_H
+#endif // SOLVERGA_H_

### 2015-06-07 include/SolverMC.h

@@ -4,7 +4,6 @@
 #include <algorithm>
 #include <vector>
 
-
 #include <assert.h>
 #include <math.h>
 
@@ -19,7 +18,8 @@
 class SolverMC {
 public:
     // constructors and destructors
-    SolverMC(// data
+    SolverMC(
+             // data
              std::vector< std::vector<double> > distance_matrix,
              std::vector<Range> ranges,
 

### 2015-06-07 include/Tree.h

@@ -42,16 +42,6 @@ private:
     std::vector<int> keyroots;
 };
 
-/*
-this distance is exactly the same as Zhang-Shasha
-tree edit distance, but based on the special case of
-unit cost for deletion and insertion and zero cost for
-substitution
-
-for RNA, this is equivalent to the number of base pairs
-to create or break to convert tree1 into tree2
-*/
-
 int unit_distance(const Tree A, const Tree B);
 
 #endif // SRC_CPP_TREE_H_

### 2015-06-07 src/RNA.cpp

@@ -1,9 +1,6 @@
 #include "../include/RNA.h"
 
 
-#define MAX_STRING_LENGTH 500
-
-
 bool is_valid_dot_bracket(std::string dot_bracket)
 {
     // tests Vienna dot-bracket for illegal structure (or symbol)
@@ -52,27 +49,29 @@ std::string only_paired(std::string dot_bracket)
 
 int levenshtein(std::string a, std::string b)
 {
-    int levDistances[MAX_STRING_LENGTH][MAX_STRING_LENGTH];
-    int aLength = 0;
-    int bLength = 0;
+    // string edit distance
+    int aLength = a.size();
+    int bLength = b.size();
+
+    std::vector< std::vector<int> > levDistances = std::vector< std::vector<int> >();
+    for(int i = 0; i < aLength+1; ++i)
+    {
+        levDistances.push_back(std::vector<int>(bLength+1, 0));
+    }
+
+
     // fill first column
-    for (int i = 0; i < MAX_STRING_LENGTH; ++i)
+    for (int i = 0; i < aLength+1; ++i)
     {
         levDistances[i][0] = i;
-        if (a[i])
-            ++aLength;
-        else
-            break;
     }
+
     // fill second column
-    for (int j = 0; j < MAX_STRING_LENGTH; ++j)
+    for (int j = 0; j < bLength+1; ++j)
     {
         levDistances[0][j] = j;
-        if (b[j])
-            ++bLength;
-        else
-            break;
     }
+
     // fill the table now
     for (int i = 0; i < aLength; ++i)
     {
@@ -127,4 +126,3 @@ Node* dot_bracket_to_node(std::string dot_bracket)
     }
     return root;
 }
-

### 2015-06-07 src/Solution.cpp

@@ -33,7 +33,7 @@ Solution::Solution(const Solution& other)
 Solution::~Solution() { }
 
 
-std::vector<int> Solution::get_genes() const
+const std::vector<int> & Solution::get_genes() const
 {
     return this->genes;
 }

### 2015-06-07 src/MCCONS_MC.cpp
changed naming for monte carlo parameter -n

@@ -95,7 +95,7 @@ int main(int argc, char *argv[])
 {
     OptionParser parser = OptionParser().description("MC-Cons Consensus Optimizer Using a Monte Carlo and Steepest Descent");
     parser.add_option("-f", "--data").dest("data_file").help("path to MARNA-like input file");
-    parser.add_option("-p", "--popsize").dest("num_trials").help("number of trials").type("size_t");
+    parser.add_option("-n", "--popsize").dest("num_trials").help("number of trials").type("size_t");
     parser.add_option("-s", "--silent").action("store_true").dest("silent").help("don't display status to stderr");
     optparse::Values options = parser.parse_args(argc, argv);
     std::vector<std::string> args = parser.args();

### 2015-06-11 doc/article.tex
added some article stuff and started Exact solver

@@
 {\Large
-\textbf\newline{Multiple RNA structure consensus assignments}
+\textbf\newline{MC-Cons: RNA secondary structure assignments}
 }
@@
 
-% example of author summary vs abstract forom Week's latest paper
-%Abstract
-%
+\section*{Abstract}
 %Discovery and characterization of functional RNA structures remains challenging due to deficiencies in de novo secondary structure modeling.
@@
 As high-quality RNA probing data become widely available, structurally-informed sequence alignment will become broadly useful for de novo motif and function discovery.
-%
-%Author Summary
-%
+RNA secondary structural assignment is an interesting problem. Bla bla bla.
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.
+
+\section*{Author Summary}
 %Despite the clear functional importance of structure in RNA molecules, it remains very difficult to correctly identify and annotate similar RNA structures because their sequences are often poorly conserved even for RNAs that form very similar higher-order structures.
@@
 We anticipate that structure-informed sequence alignment and structure modeling will become broadly useful tools in RNA function analysis.
-
-\section*{Abstract}
-RNA secondary structural assignment is an interesting problem. Bla bla bla.
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.
-
-\section*{Author Summary}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
@@
 \section*{Materials and Methods}
-\subsection*{Theoretical foundation}
+\subsection*{Theoretical foundation: location theory}
+The idea of minimizing the sum of pairwise distance between objects is not new.
+This has been investigated in location theory, with the goal of assigning position
+to factories for example.
 
@@
 
-\subsection*{MCCONS}
+\subsection*{MC-Cons}
 
-\paragraph{Metrics}
+\paragraph{Distance functions}
 

### 2015-06-11 include/SolverExact.h

@@ -1,0 +1,31 @@
+#ifndef SOLVEREXACT_H_
+#define SOLVEREXACT_H_
+
+#include "Solver.h"  // abstract class and useful functions
+
+#include <assert.h>
+
+
+class SolverExact: public Solver{
+public:
+    // constructors and destructors
+    SolverExact(// misc parameters
+                bool silent,
+                // exact solver parameters
+                double best_known_score);
+    ~SolverExact();
+
+    // solver call
+    std::vector<Solution> solve(std::vector< std::vector<double> > distance_matrix,
+                                std::vector<Range> ranges,
+                                unsigned long seeds[6]) const;
+    void set_threshold(double new_threshold);
+
+private:
+    double threshold_;
+};
+
+
+
+
+#endif // SOLVEREXACT_H_

### 2015-06-11 src/MCCONS_EXACT.cpp

@@ -1,187 +1,48 @@
-#include "../include/RNA.h"
-#include "../include/RNAConsensus.h"
-#include "../include/SolverGA.h"
-#include "../include/Solution.h"
 
 
-using std::vector;
-using std::string;
+#include "../include/SolverExact.h"
+#include "../include/OptionParser.h"
 
 
-double calculate_insertion_cost(std::vector<int> selected_indices,
-                                int elem_index,
-                                std::vector< std::vector<double> > distance_matrix)
+using optparse::OptionParser;
+
+
+int main(int argc, char *argv[])
 {
-    // calculate the cost of adding an element into the solution
-    double cost = 0.;
-    int index;
-    for (size_t i = 0; i < selected_indices.size(); ++i)
+
+    // create the command line parser
+    OptionParser parser = OptionParser().description("MC-Cons Consensus Optimizer Using an Exact Branch and Bound Solver");
+    parser.add_option("-f", "--data").dest("data_file").help("path to MARNA-like input file");
+    parser.add_option("-h", "--higher_bound").dest("higher_bound").help("best score known").type("size_t");
+    parser.add_option("-s", "--silent").action("store_true").dest("silent").help("don't display status to stderr");
+    optparse::Values options = parser.parse_args(argc, argv);
+    std::vector<std::string> args = parser.args();
+
+
+    if (options.is_set("data_file") && options.is_set("higher_bound"))
     {
-        index = selected_indices[i];
-        cost += distance_matrix[index][elem_index];
-    }
-    return (2. * cost);  // symmetric
-}
+      // extract command line parameters
+      bool silent = false;
+      if (options["silent"] == "1")
+      {
+          silent = true;
+      }
 
+      std::string path = options["data_file"];
+      size_t population_size = atoi(options["higher_bound"].c_str());
+      size_t num_generations = atoi(options["numGenerations"].c_str());
 
-struct InsertionPoints {std::vector<int> best_insertions, std::vector<double> best_costs};
+      // instantiate the genetic algorithm solver
+      Solver* solver = new SolverExact(silent,
+                                       best_known_score);
 
-
-InsertionPoints find_best_insertion(std::vector<int> selected_indices,
-                                    std::vector<Range> remaining_ranges,
-                                    std::vector< std::vector<double> > distance_matrix)
-{
[…]

### 2015-06-11 src/SolverExact.cpp

@@ -1,0 +1,24 @@
+
+
+// constructors and destructors
+SolverExact::SolverExact(bool silent) : Solver(silent)
+{
+    threshold_ = std::numeric_limits<double>::infinity();
+}
+
+
+SolverExact::~SolverExact() { }
+
+
+SolverExact::set_threshold(double new_threshold)
+{
+    threshold_ = new_threshold;
+}
+
+
+std::vector<Solution> SolverExact::solve(std::vector< std::vector<double> > distance_matrix,
+                                         std::vector<Range> ranges,
+                                         unsigned long seeds[6]) const
+{
+    
+}

### 2015-06-11 src/SolverGA.cpp

@@ -4,24 +4,19 @@
 using std::vector;
 
 
-
-
 // constructors and destructors
-SolverGA::SolverGA( // data
-                    // misc parameters
-                    bool silent,
+SolverGA::SolverGA(// misc parameters
+                   bool silent,
 
-                    // GA settings
-                    int population_size,
-                    int num_generations,
-                    int improvement_depth,
-                    int elite_size,
+                   // GA settings
+                   int population_size,
+                   int num_generations,
+                   int improvement_depth,
+                   int elite_size,
 
-                    double crossover_prob,
-                    double mutation_prob,
-                    double improvement_prob
-
-                    ) : Solver(silent)
+                   double crossover_prob,
+                   double mutation_prob,
+                   double improvement_prob) : Solver(silent)
 {
 
     population_size_ = population_size;

### 2015-06-11 src/SolverMC.cpp

@@ -83,5 +83,6 @@ std::vector<Solution> SolverMC::solve(std::vector< std::vector<double> > distanc
         bar.clean();
     }
 
+    delete prng;
     return best_solutions;
 }

### 2015-06-11 src/cluster.py
added sequences and python clustering

@@ -1,0 +1,356 @@
+# some clustering utilities
+import scipy.cluster
+import seaborn
+
+
+def get_dist_matrix(sol, problem):
+    """creates a new distance matrix with only solutions"""
+    matrix_size = len(sol.genes)
+    distance_matrix = np.zeros((matrix_size, matrix_size))
+    for index1, gene1 in enumerate(sol.genes):
+        for index2, gene2 in enumerate(sol.genes):
+            if index2 > index1:
+                dist = problem.distance_matrix[gene1][gene2]
+                distance_matrix[index1][index2] = dist
+                distance_matrix[index2][index1] = dist
+    return distance_matrix
+
+
+def get_cluster_information(distance_matrix, cluster_method="average"):
+    """returns the order in which the consensus should be presented"""
+    # perform the hierarchical clustering
+    mat_ut = distance_matrix[np.triu_indices(distance_matrix.shape[0], 1)]
+    Z = scipy.cluster.hierarchy.linkage(mat_ut, method=cluster_method)
+    
+    # convert the linkage information to a dendogram ordering
+    # http://stackoverflow.com/questions/12572436/calculate-ordering-of-dendrogram-leaves
+    n = len(Z) + 1
+    cache = dict()
+    for k in range(len(Z)):
+        c1, c2 = int(Z[k][0]), int(Z[k][1])
+        c1 = [c1] if c1 < n else cache.pop(c1)
+        c2 = [c2] if c2 < n else cache.pop(c2)
+        cache[n+k] = c1 + c2
+    return cache[2*len(Z)], Z
+
+
+def show_solution(solution, problem, cluster_method="average", show_objects=True):
+    """show a dendrogram with the distance matrix"""
+    distance_matrix = get_dist_matrix(solution, problem)
+    print_order, linkage = get_cluster_information(distance_matrix, cluster_method=cluster_method)
+    
+    # get the objects
+    if show_objects:
+        ylabels = []
+        max_num_digits = divmod(len(print_order), 10)[0]
+        for position in print_order:
+            gene = solution.genes[position]
+            num_digits = divmod(position, 10)[0]
+            padding = "".join([" " for _ in range(max_num_digits - num_digits)])
+            ylabels.append("[{}] {} {}".format(position, padding, problem.objects[gene]))
+        return clustermap(distance_matrix, row_linkage=linkage, annot=True, col_cluster=False, yticklabels=ylabels, cbar=False)
+    else:
[…]

### 2015-06-19 article/article.tex
started working more on the article, added lots of refs

--- 2015-06-11 doc/article.tex
@@
 \newpage
-\section*{Materials and Methods}
-\subsection*{Theoretical foundation: location theory}
-The idea of minimizing the sum of pairwise distance between objects is not new.
-This has been investigated in location theory, with the goal of assigning position
-to factories for example.
+\section*{Methods}
 
-\subsection*{Metaheuristics}
+\subsection*{Simple Tree Distance for Secondary Structure}
+
+\paragraph{Base Pair Tree Representation}
+% base pair tree representation vs abstract shapes
+% upsides and downsides of the
+% representation compared to abstract shape
+RNA secondary structure can represented by base pair trees, where only base pairs are considered.
+This representation is very similar to level 3 RNA abstract shape \cite{abstract_shapes}, albeit the stem length information is conserved.
+Although this representation greatly increases the size of the auxiliary space (codomain) compared to level 3 abstract shapes, it doesn't suffer from major changes in representation if small stem-loops are added or removed.
+It also doesn't map structures of widely varying stem length to the same element in the codomain.
+
+%example of the conversion process 
+\noindent The mapping from a Vienna dot bracket representation can be made simply by removing the dots (unpaired nucleotides) and then creating a tree from it.
+Since all nodes represent base pairs, the resulting tree can be considered unlabelled.
+Note that an artificial root is added to all trees.
+
+\paragraph{Unit Tree Indel Distance}
+% relation to the tree edit distance (special scoring scheme)
+The unit tree indel distance is a simple and intuitive distance function based on the well known tree edit distance \cite{tai, zhang_shasha}, where insertions and deletions have a unit cost.
+Since it is applied to base pair trees which are unlabelled, node relabelling comes at no cost.
+
+% interpretation of a score
+\noindent Using this scoring scheme, a distance of $n$ means that the smallest number of base pair breakage or formation to transition from a base pair tree to another is $n$.
+The distance function retains the metric properties.
+
+\subsection*{Optimization goal}
+
+The idea of minimizing the sum of pairwise distance between sets of objects if far from new and a significant body of work has been done in the field of location theory \cite{compact_location}.
 
@@
 
-\paragraph{Distance functions}
+MC-Cons is a two-step optimization process relying on the minimization of the sum of pairwise distance of two distance functions. First, the
 
-\subparagraph{Tree Indel distance on base pair tree}
-
-RNA secondary structure is represented as a tree, where nodes are base pairs and unpaired positions are ignored.
-
-\subparagraph{String Edit Distance}
-%\subsection*{}
+\subsection*{Metaheuristics}
 
@@
 
-%\section*{References}
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-% OR
-%
+\section*{References}
 % Compile your BiBTeX database using our plos2015.bst
@@
 
+\begin{thebibliography}{10}
+\bibitem{abstract_shapes}
+Giegerich, Robert and Vo{\ss}, Bj\"{o}rn and Rehmsmeier, Marc. Abstract shapes
+    of RNA. Nucleic Acids Research, 32(16):4843–4851, January 2004.
+
+\bibitem{tai}
+Kuo-Chung Tai. 1979. The Tree-to-Tree Correction Problem. J. ACM 26, 3 (1979), 422-433.
+
+\bibitem{zhang_shasha}
+Dennis Shasha and Kaizhong Zhang. Fast algorithms for the unit cost
+    editing distance between trees. Journal of Algorithms, 11(4):581–621,
+    December 1990.
+
+\bibitem{compact_location}
+ S. O. Krumke, M. V. Marathe, H. Noltemeier, et al. Compact location problems Theoretical Computer Science, Vol. 181, No. 2. (1997), pp. 379-404
+\end{thebibliography}
+
 \end{document}

### 2015-06-21 article/article.tex
saving for today and added article on scientific writing

@@
 
+% WHAT KIND OF ARTICLE? : software paper right? although it's more algorithm with implementation given
+
 \section*{Abstract}
@@
 
-\subsection*{Simple Tree Distance for Secondary Structure}
-
-\paragraph{Base Pair Tree Representation}
+\subsection*{Base Pair Tree Representation of Secondary Structure}
 % base pair tree representation vs abstract shapes
@@
 RNA secondary structure can represented by base pair trees, where only base pairs are considered.
-This representation is very similar to level 3 RNA abstract shape \cite{abstract_shapes}, albeit the stem length information is conserved.
-Although this representation greatly increases the size of the auxiliary space (codomain) compared to level 3 abstract shapes, it doesn't suffer from major changes in representation if small stem-loops are added or removed.
-It also doesn't map structures of widely varying stem length to the same element in the codomain.
+This representation is very similar to level 5 RNA abstract shape \cite{abstract_shapes}, albeit the stem length information is conserved.
+Although this representation greatly increases the size of the auxiliary space (codomain) compared to level 5 abstract shapes, it does not suffer from major changes in representation if small stem-loops are added or removed.
+It also does not map structures of widely varying stem length to the same element in the codomain.
 
@@
 Since all nodes represent base pairs, the resulting tree can be considered unlabelled.
-Note that an artificial root is added to all trees.
+Note that an artificial root is added to all structures.
 
-\paragraph{Unit Tree Indel Distance}
+\subsection*{Unit Tree Indel Distance}
+
 % relation to the tree edit distance (special scoring scheme)
@@
 % interpretation of a score
-\noindent Using this scoring scheme, a distance of $n$ means that the smallest number of base pair breakage or formation to transition from a base pair tree to another is $n$.
+\noindent Under this scoring scheme, a distance of $n$ means that the smallest number of base pair breakage or formation to transition from a base pair tree to another is $n$.
 The distance function retains the metric properties.
 
-\subsection*{Optimization goal}
+\noindent Overall, the unit tree indel distance is an interesting distance function with both intuitive interpretation and very useful mathematical properties.
 
-The idea of minimizing the sum of pairwise distance between sets of objects if far from new and a significant body of work has been done in the field of location theory \cite{compact_location}.
+\subsection*{Minimizing Sum of Pairwise Distances}
+
+% introduce the problem
+
+% sp-alignment
+
+% introduce compact location problem
+
+% basic properties
+%\noindent Naively, given $n$ molecules each with $m$ secondary structures, the going through the full search space would involve the evaluation of $m^n$ solutions.
+This is obviously not a good ideaHowever, the idea of
+
+Minimizing the sum of pairwise distance between sets of objects is far from new and a significant body of work has been done in the field of location theory \cite{compact_location}. \newline
+
+% mapping of the consensuss problem to the compact location problem by
+% tweaking the distance function using infinity as intra group distance
+% may need more details to proove it, let's see what Francois thinks
+\noindent The consensus problem described here can be reduced to an instance of the minimum average distance placement (MIN-SUM) by modifying the distance function in use such that elements belonging to the same group have infinite distance, thereby guaranteeing that a consensus will never contain more than one element of each group.
+This convenient reduction allows us to benefit from previously derived hardness results and justify the use of meta-heuristics.
+
+\noindent Using a metric distance function, one might think that it would be possible to reduce efficiently the search space, but even finding a near-optimal placement within a certain factor of the optimal has been proved to be NP-Hard \cite{compact_location}.
+
+\subsection*{Heuristic Consensus Solvers}
+% complexity justifies meta-heuristic
+Given the complexity of the consensus problem, the use of heuristics is well justified.
+
+% check which details should be given, keep it as simple as possible
+% article using metaheuristics? stuff from class maybe?
+
+% mention the interface, the solvers are given a precomputed distance matrix!
+\noindent Many simple heuristics were found to perform well within reasonable time. These algorithms all take precomputed distance matrices whose calculation is embarrassingly parallel.
+
+% for the article, we'll ignore the steepest descent solver for now, it is useless
+% version 1: Exact Version (branch & bound)
+\paragraph{Exact Version: Branch \& Bound}
+First, an exact version of the consensus solver based on a branch \& bound design was created to compare the performance of different heuristics on small instances.
+
+\noindent The lower bound of an incomplete solution is estimated by the summation of pairwise distances between the chosen objects and the best distance they have to remaining sets of objects to be chosen.
+This bound is generally quite good on small instances, but its performance degrades rapidly as the number of molecules grows.
+
+% version 2: Genetic Algorithm
+\paragraph{Heuristic: Genetic Algorithm}
+The genetic algorithm solver uses generic mutation (random substitution) and crossover operators (uniform).
+To increase convergence rate, an improvement operator using a steepest descent procedure is used.
+
+\noindent The local search is done by finding the substitution which improves the most the current sum of pairs distance.
+The operator applies a small number of iterations of local search with a certain probability.
 
 \subsection*{MC-Cons}
+% specify that this is an heuristic
 
-MC-Cons is a two-step optimization process relying on the minimization of the sum of pairwise distance of two distance functions. First, the
+MC-Cons is a two-step optimization procedure relying on the minimization of the sum of pairwise distance of two distance functions to create a structural assignment.
 
-\subsection*{Metaheuristics}
+% minimizing the tree distance
+\noindent First, \textit{tree consensus} are calculated based on the base pair tree representation of the secondary structures.
+The intuition is that the topology of the tree is the single most important aspect of structural comparison.
+The unit tree indel distance
+
+% within tree consensus, minimizing a naive distance: string edit distance
+\noindent For each established tree consensus, all secondary structures whose base pair tree representation is in the tree consensus is filtered and a string edit distance consensus is calculated.
 
@@
 
-% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}. 
-\subsection*{IRES}
+% Please do not create a heading level below \subsection. For 3rd level headings, use \subsubsection*{}. 
+\subsection*{Basic Example: IRES}
+
+Iron Response Elements (IREs) can be used as a toy example.
+
+Take the 10 first suboptimals
 
@@
 For more information, see \nameref{S1_Text}.
+
+\section*{Software Availability} %taken from dcGOR: Software for Ontologies and Domain Annotation
+As open-source software, MC-Cons is freely available under the GNU LGPL v3.0 license.
+It is distributed at GitHub, https://github.com/major-lab/MC-Cons along with documentation and working examples.
+The algorithms are implemented in C++ with no external dependencies and offers a very simple command line interface.
 

### 2015-06-21 src/MCCONS_GA.cpp
set default parameters for all tools, only specify file now

@@ -30,25 +30,36 @@ int main(int argc, char *argv[])
     double CROSSOVER_PROBABILITY = 0.5;
     double MUTATION_PROBABILITY = 0.05;
     double IMPROVEMENT_PROBABILITY = 0.05;
+    size_t POPULATION_SIZE = 250;
+    size_t NUM_GENERATIONS = 250;
+    bool silent = false;
 
 
-    if (options.is_set("dataFile") && options.is_set("popSize") && options.is_set("numGenerations"))
+    if (options.is_set("dataFile"))
     {
       // extract command line parameters
-      bool silent = false;
       if (options["silent"] == "1")
       {
           silent = true;
       }
 
+      if (options.is_set("popSize"))
+      {
+          POPULATION_SIZE = atoi(options["popSize"].c_str());
+      }
+
+      if (options.is_set("numGenerations"))
+      {
+          NUM_GENERATIONS = atoi(options["numGenerations"].c_str());
+      }
+
+
       std::string path = options["dataFile"];
-      size_t population_size = atoi(options["popSize"].c_str());
-      size_t num_generations = atoi(options["numGenerations"].c_str());
 
       // instantiate the genetic algorithm solver
       Solver* solver = new SolverGA(silent,
-                                    population_size,
-                                    num_generations,
+                                    POPULATION_SIZE,
+                                    NUM_GENERATIONS,
                                     IMPROVEMENT_DEPTH,
                                     ELITE_SIZE,
 

### 2015-06-21 src/MCCONS_MC.cpp

@@ -16,9 +16,10 @@ int main(int argc, char *argv[])
 
     // GLOBAL SETTINGS (MODIFY AT YOUR OWN RISKS (WHICH ARE MINIMAL))
     unsigned long SEEDS[6] = {42, 42, 42, 42, 42, 42};
+    size_t NUM_TRIALS = 500;
 
     // extract parameters and lauch the genetic algorithm (either silent or not)
-    if (options.is_set("data_file") && options.is_set("num_trials"))
+    if (options.is_set("data_file"))
     {
       // extract command line parameters
       bool silent = false;
@@ -27,10 +28,13 @@ int main(int argc, char *argv[])
           silent = true;
       }
       std::string path = options["data_file"];
-      size_t num_trials = atoi(options["num_trials"].c_str());
+      if (options.is_set("num_trials"))
+      {
+          NUM_TRIALS = atoi(options["num_trials"].c_str());
+      }
 
       // instantiate the genetic algorithm solver
-      Solver* solver = new SolverMC(num_trials, silent);
+      Solver* solver = new SolverMC(NUM_TRIALS, silent);
 
       // execute and cleanup
       MCCONS(path, solver, SEEDS);

### 2015-06-21 article/article.tex
got a first version of the method section

@@
 
-\subsection*{Heuristic Consensus Solvers}
+\subsection*{Consensus Solvers}
 % complexity justifies meta-heuristic
@@
 \paragraph{Heuristic: Genetic Algorithm}
-The genetic algorithm solver uses generic mutation (random substitution) and crossover operators (uniform).
+The genetic algorithm solver uses generic mutation (random substitution) and crossover operators (uniform) along with elitist selection.
 To increase convergence rate, an improvement operator using a steepest descent procedure is used.
@@
 \subsection*{MC-Cons}
-% specify that this is an heuristic
-
-MC-Cons is a two-step optimization procedure relying on the minimization of the sum of pairwise distance of two distance functions to create a structural assignment.
-
-% minimizing the tree distance
-\noindent First, \textit{tree consensus} are calculated based on the base pair tree representation of the secondary structures.
-The intuition is that the topology of the tree is the single most important aspect of structural comparison.
-The unit tree indel distance
 
 % within tree consensus, minimizing a naive distance: string edit distance
-\noindent For each established tree consensus, all secondary structures whose base pair tree representation is in the tree consensus is filtered and a string edit distance consensus is calculated.
+MC-Cons is a two-step optimization procedure (Fig.~\ref{fig1}).
+First, \textit{base pair tree consensus} are calculated.
+This insures that the final consensus will minimize the difference in base pair topology.
+For each base pair tree consensus, the subset of secondary structures whose base pair tree are  in the consensus are kept.
+String edit distance consensus (on Vienna dot brackets) are then calculated.
+Normally this distance function would not be suitable, but since tree distance is already minimized, the final consensus is satisfying.\newline
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[scale=0.4]{figs/mccons_flowchart2.jpg}
+\caption{{\bf Flowchart illustration of MC-Cons procedure.}
+The two-step procedure is quite simple. The user only inputs data and chooses which consensus solver, either the exact one for small instances or the genetic algorithm for larger problems.}
+\label{fig1}
+\end{center}
+\end{figure}
 
@@
 % For figure citations, please use "Fig." instead of "Figure".
-Nulla mi mi, Fig.~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam.
+%Nulla mi mi, Fig.~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam.
 Proin rutrum vel massa non gravida.
@@
 Praesent faucibus semper libero.
-
-\begin{figure}[h]
-\caption{{\bf Figure Title first bold sentence Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}
-Figure Caption Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
-\label{fig1}
-\end{figure}
 

### 2015-06-21 article/article.tex
removed jpg images, replace by .eps and refactored text

@@
 This insures that the final consensus will minimize the difference in base pair topology.
-For each base pair tree consensus, the subset of secondary structures whose base pair tree are  in the consensus are kept.
+For each base pair tree consensus and for each group of structures, the subset of secondary structures whose base pair tree fits the one assigned to that group is kept.
 String edit distance consensus (on Vienna dot brackets) are then calculated.
-Normally this distance function would not be suitable, but since tree distance is already minimized, the final consensus is satisfying.\newline
+Normally this distance function would not be sufficient, but since tree distance is already minimized, the final consensus is satisfactory.\newline
 
@@
 \begin{center}
-\includegraphics[scale=0.4]{figs/mccons_flowchart2.jpg}
+\includegraphics[scale=0.4]{figs/mccons_flowchart2.eps}
 \caption{{\bf Flowchart illustration of MC-Cons procedure.}

### 2015-06-22 include/Solution.h
small refactoring of Solution class

@@ -1,17 +1,16 @@
 #ifndef SOLUTION_H
 #define SOLUTION_H
 
-#include <iostream>
+#include <iostream>  // <<
 #include <vector>
-#include <limits>
+#include <limits>    // infinity for default score
 
 class Solution
 {
 public:
     // constructors and destructors
-    Solution();
-    Solution(std::vector<int> chosen_genes);
-    Solution(std::vector<int> chosen_genes, double score);
+    Solution(std::vector<int> genes=std::vector<int>(),
+             double score=std::numeric_limits<double>::infinity());
     Solution(const Solution& other); // copy constructor
 
     ~Solution();
@@ -19,7 +18,6 @@ public:
     // getters and setters
     const std::vector<int> & get_genes() const;
     void set_gene(int position, int new_gene);
-    void add_gene(int new_gene);
 
     double get_score() const;
     void set_score(double new_score);
@@ -32,8 +30,8 @@ public:
     friend std::ostream& operator<< (std::ostream &out, Solution &sol);
 
 private:
-    std::vector<int> genes;
-    double score;
+    std::vector<int> genes_;
+    double score_;
 };
 
 

### 2015-06-22 src/Solution.cpp

@@ -1,32 +1,19 @@
 #include "../include/Solution.h"
 
 
-Solution::Solution()
+Solution::Solution(std::vector<int> genes, double score)
 {
-    genes = std::vector<int>();
-    score = std::numeric_limits<double>::infinity();
+    genes_ = genes;
+    score_ = score;
+
 }
 
-
-Solution::Solution(std::vector<int> chosen_genes)
-{
-    genes = chosen_genes;
-    score = std::numeric_limits<double>::infinity();
-}
-
-
-Solution::Solution(std::vector<int> chosen_genes, double calculated_score)
-{
-    this->genes = chosen_genes;
-    this->score = calculated_score;
-
-}
 
 Solution::Solution(const Solution& other)
 {
     // copy constructor
-    genes = other.get_genes();
-    score = other.get_score();
+    genes_ = other.get_genes();
+    score_ = other.get_score();
 }
 
 
@@ -35,40 +22,35 @@ Solution::~Solution() { }
 
 const std::vector<int> & Solution::get_genes() const
 {
-    return this->genes;
+    return genes_;
 }
+
 
 void Solution::set_gene(int position, int new_gene)
 {
-    this->genes[position] = new_gene;
-}
-
-
-void Solution::add_gene(int new_gene)
-{
-    this->genes.push_back(new_gene);
+    genes_[position] = new_gene;
 }
 
 
 double Solution::get_score() const
 {
-    return this->score;
+    return score_;
 }
 
 void Solution::set_score(double new_score)
 {
-    this->score = new_score;
+    score_ = new_score;
 }
 
 bool Solution::operator<(const Solution &other) const
 {
-    return this->score < other.score;
+    return score_ < other.score_;
 }
 
 
 bool Solution::operator==(const Solution &other)
 {
-    return this->genes == other.genes;
+    return genes_ == other.genes_;
 }
 
 
@@ -76,12 +58,11 @@ std::ostream& operator<< (std::ostream &out, Solution &sol)
 {
     // outputs the genes and the score to the stream
     out << "[";
-    for (std::vector<int>::iterator it = sol.genes.begin();
-            it != sol.genes.end(); ++it)
+    for (size_t i = 0; i != sol.genes_.size(); ++i)
     {
-        out << *it << " ";
+        out << sol.genes_[i] << " ";
     }
-    out << " : " << sol.score << "]" << std::endl;
+    out << " : " << sol.score_ << "]" << std::endl;
     return out;
 }
 

### 2015-06-22 include/ConsensusProblem.h
did some minor code refactoring

@@ -12,7 +12,7 @@ struct Range
 {
     int low;
     int high;
-}; // to keep the range
+}; // to keep the boundaries of each group of objects
 
 
 template <class T>
@@ -36,6 +36,8 @@ private:
     std::vector< std::vector<double> > distance_matrix;
 };
 
+// used to assert that the distance matrix passed
+// is at least symmetric and has zero on the diagonal
 bool distance_is_symmetric(const std::vector< std::vector<double> > & distance_matrix);
 
 

### 2015-06-22 include/MCCONS.h

@@ -4,6 +4,8 @@
 #include "Solver.h"
 #include "RNAConsensus.h"
 
+// The whole two-step optimization procedure is done by this
+// function which is controlled by feeding different solvers
 void MCCONS(std::string path,
             Solver* solver,
             unsigned long seeds[6]);

### 2015-06-22 include/Node.h

@@ -31,10 +31,10 @@ private:
 };
 
 
-// node <-> bracket interconversion
+// node -> bracket reduction
 void node_to_bracket(Node* node, std::vector<char>& brackets);
 
-// helpers
+// computes the postorder enumeration of the nodes
 std::vector<Node*> get_postorder_enumeration(Node* root);
 
 

### 2015-06-22 include/ProgressBar.h

@@ -16,12 +16,12 @@ public:
     // updates the progress displayed
     void update(float percentage);
 
-    // displays 100% status
+    // displays 100% status, should be called immediately after last iteration
     void clean();
 
 private:
-    size_t _bar_length;
-    std::string _process_name;
+    size_t bar_length_;
+    std::string process_name_;
 };
 
 #endif // PROGRESSBAR_H

### 2015-06-22 include/RNA.h

@@ -1,20 +1,0 @@
-#ifndef SRC_CPP_RNA_H_
-#define SRC_CPP_RNA_H_
-
-#include <algorithm>
-
-#include "Tree.h"
-#include "RngStream.h"
-
-
-// rna secondary structure general helpers (conversion and distances)
-
-bool is_valid_dot_bracket(std::string dot_bracket);
-
-std::string only_paired(std::string dot_bracket);
-
-int levenshtein(std::string a, std::string b);
-
-Node* dot_bracket_to_node(std::string  dot_bracket);  // don't forget to free it
-
-#endif  // SRC_CPP_RNA_H_

### 2015-06-22 include/RNAConsensus.h

@@ -1,17 +1,27 @@
 #ifndef RNACONSENSUS_H
 #define RNACONSENSUS_H
 
-#include "RNA.h"  // need some tree and dot bracket methods
+#include "Tree.h"
+
+#include <algorithm>
 #include <string>
 #include <iostream> // file reading
-#include <fstream>
+#include <fstream>  // ifstream
 #include <ctype.h>  // tolower()
 #include <stdlib.h> // for exit
 
+// rna secondary structure general helpers (conversion and distances)
+
+
+int levenshtein(std::string a, std::string b);
+
+
+// file IO and MC-Cons utilities
 std::vector<std::vector<std::string> > read_data(std::string file_name);
 
 std::vector< std::vector<Tree> > get_tree_lists(std::vector< std::vector<std::string> > dot_brackets_vectors);
 
 std::vector< std::vector<std::string> > filter_dot_brackets(std::vector< std::vector<std::string> > dot_brackets_vectors, std::vector<std::string> brackets);
+
 
 #endif // RNACONSENSUS_H

### 2015-06-22 include/Solver.h

@@ -14,7 +14,8 @@
 #include <vector>
 
 // interface for all solvers
-
+// any new solver must inherit this class if you
+// want to plug it easily in the MC-Cons algorithm
 class Solver
 {
 public:
@@ -22,6 +23,9 @@ public:
     {
         silent_ = silent;
     }
+
+    // used by MC-Cons, accesses the status of the solver
+    bool is_silent();
 
     virtual std::vector<Solution> solve(std::vector< std::vector<double> > distance_matrix,
                                         std::vector<Range> ranges,
@@ -40,19 +44,26 @@ struct int_double
 };
 
 
+// calculates and assigns the sum of pairwise distances
+// of a solution and assign it as score_
 void assign_pairwise_distance_score(Solution& sol,
                                     std::vector< std::vector<double> > distance_matrix);
 
 
+// randomly chooses numbers fitting the boundaries of the ranges
+std::vector<int> select_random_genes(std::vector<Range> ranges, RngStream *prng);
+
+// create a list of Solutions generated randomly
 std::vector<Solution> initialize_population(std::vector<Range> ranges,
                                             int pop_size, RngStream *prng);
 
 
+// applies the steepest descent procedure until either
+// a local minima is found or max_iter iterations of the substitution are done
 void steepest_improvement(Solution& sol,
                           std::vector< std::vector<double> > distance_matrix,
                           std::vector<Range> ranges,
                           int max_iter=std::numeric_limits<int>::infinity());
 
-std::vector<int> select_random_genes(std::vector<Range> ranges, RngStream *prng);
 
 #endif  // SOLVER_H

### 2015-06-22 include/SolverExact.h

@@ -3,7 +3,7 @@
 
 #include "Solver.h"  // abstract class and useful functions
 #include "Solution.h"
-#include "SolutionNode.h"
+#include "SubSolution.h"
 
 #include <algorithm>
 #include <limits>
@@ -11,11 +11,14 @@
 
 #include <assert.h>
 
+
+// solves the consensus problem using a branch and bound design
+// calculation of the lower bound of a subsolution is described in the
+// Solver.cpp implementation file
 class SolverExact: public Solver{
 public:
     // constructors and destructors
     SolverExact(bool silent);
-
     ~SolverExact();
 
     // solver call

### 2015-06-22 include/SubSolution.h

@@ -17,20 +17,19 @@ struct RangeIndex
 
 
 // special solution type used for exact solver
-class SolutionNode
+class SubSolution
 {
 public:
     // constructor and destructor
-    SolutionNode(std::vector<int> genes,
+    SubSolution(std::vector<int> genes,
                  std::vector<RangeIndex> ranges,
                  double score = std::numeric_limits<double>::max());
-    SolutionNode(const SolutionNode& other);
-    ~SolutionNode();
+    SubSolution(const SubSolution& other);
+    ~SubSolution();
 
     // getters and setters
     const std::vector<int> & get_genes() const;
     void set_gene(int position, int new_gene);
-    void add_gene(int new_gene);
 
     std::vector<RangeIndex> get_ranges() const;
 
@@ -42,11 +41,11 @@ public:
     Solution return_solution();
 
     // some comparison operators
-    bool operator <(const SolutionNode &other) const;
-    bool operator ==(const SolutionNode &other);
+    bool operator <(const SubSolution &other) const;
+    bool operator ==(const SubSolution &other);
 
     // pretty printer
-    friend std::ostream& operator<< (std::ostream &out, const SolutionNode &sol);
+    friend std::ostream& operator<< (std::ostream &out, const SubSolution &sol);
 
 
 private:

### 2015-06-22 include/Tree.h

@@ -9,7 +9,6 @@
 #include <assert.h>
 
 #include "Node.h"
-#include "RNA.h"
 
 
 class Tree
@@ -41,6 +40,20 @@ private:
     std::vector<int> leftmost_descendents;
     std::vector<int> keyroots;
 };
+
+
+// used to assert that dot bracket is balanced
+// and contains no illegal symbols
+bool is_valid_dot_bracket(std::string dot_bracket);
+
+
+// (((..).))) -> ((()))
+std::string only_paired(std::string dot_bracket);
+
+
+// converts dot bracket to Nodes, used by Tree constructor
+// don't forget to free it
+Node* dot_bracket_to_node(std::string  dot_bracket);
 
 int unit_distance(const Tree A, const Tree B);
 

### 2015-06-22 src/MCCONS.cpp

@@ -7,17 +7,24 @@ void MCCONS(std::string path,
             unsigned long seeds[6])
 {
 
-    // TREE CONSENSUS
-    // fetch the data and instantiate the tree consensus problem
+    // PHASE 1: TREE CONSENSUS
+
+    // data acquisition
     std::vector<std::vector<std::string> > dot_brackets = read_data(path);
     std::vector<std::vector<Tree> > trees = get_tree_lists(dot_brackets);
     ConsensusProblem<Tree> tree_problem = ConsensusProblem<Tree>(trees, *unit_distance);
+
+    // PHASE 1 SOLVING
+    if (! solver->is_silent())
+    {
+        std::cerr << "Phase 1: Base Pair Tree Consensus (1)" << std::endl;
+    }
 
     std::vector<Solution> tree_consensus = solver->solve(tree_problem.get_distance_matrix(),
                                                          tree_problem.get_ranges(),
                                                          seeds);
 
-    // filter the dot brackets by the tree consensus
+    // PHASE 1.5: FILTERING
     std::vector<std::vector< std::vector<std::string> > > prob2_data = std::vector< std::vector< std::vector<std::string> > >();
 
     // get the brackets
@@ -39,7 +46,12 @@ void MCCONS(std::string path,
     }
 
 
-    // TREE-STRING CONSENSUS
+    // PHASE 2: STRING CONSENSUS
+    if (! solver->is_silent())
+    {
+        // update on the number of optimal trees found
+        std::cerr << std::endl << "Phase 2: String Edit Distance on Vienna Dot Bracket (" << dot_bracket_problems.size() << ")" << std::endl;
+    }
     std::vector< std::vector<Solution> > dot_bracket_consensus = std::vector< std::vector<Solution> >();
     for (size_t i = 0; i != prob2_data.size(); ++i)
     {
@@ -57,6 +69,10 @@ void MCCONS(std::string path,
                 best_dot_bracket_score = dot_bracket_consensus[i][j].get_score();
             }
         }
+    }
+    if (! solver->is_silent())
+    {
+        std::cerr << std::endl;
     }
 
     // printing the results to stdout

### 2015-06-22 src/ProgressBar.cpp

@@ -3,8 +3,8 @@
 
 ProgressBar::ProgressBar(std::string process_name, size_t bar_length)
 {
-    _process_name = process_name;
-    _bar_length = bar_length;
+    process_name_ = process_name;
+    bar_length_ = bar_length;
 }
 
 
@@ -17,22 +17,24 @@ void ProgressBar::update(float percentage)
     assert (percentage >= 0.);
     assert (percentage <= 1.);
 
-    size_t progress_length = _bar_length * percentage;
+    size_t progress_length = bar_length_ * percentage;
     std::cerr << "[";
-    for (size_t i = 0; i < _bar_length; ++i)
+    for (size_t i = 0; i < bar_length_; ++i)
     {
         if (i <= progress_length)
             std::cerr << "#";
         else
             std::cerr << " ";
     }
-    std::cerr << "] " << int(percentage * 100) << "% " << _process_name << "\r";// ;
+    std::cerr << "] " << int(percentage * 100) << "% " << process_name_ << "\r";// ;
     std::cerr.flush();
 }
 
 
 void ProgressBar::clean()
 {
+    // called after last iteration, displays 100% success
+    // because I seem to constantly mess up (usually ends 99%)
     this->update(1.);
     std::cerr << std::endl;
     std::cerr.flush();

### 2015-06-22 src/RNA.cpp

@@ -1,128 +1,0 @@
-#include "../include/RNA.h"
-
-
-bool is_valid_dot_bracket(std::string dot_bracket)
-{
-    // tests Vienna dot-bracket for illegal structure (or symbol)
-    int counter = 0;
-
-    for (unsigned int i = 0; i < dot_bracket.size(); ++i)
-    {
-        char c = dot_bracket[i];
-        if (c == '(')               // stack
-        {
-            counter += 1;
-        }
-        else if (c == ')')          // unstack
-        {
-            counter -= 1;
-        }
-        else if (c != '.')          // illegal character
-        {
-            return false;
-        }
-        if (counter < 0)            // left unbalanced
-        {
-            return false;
-        }
-    }
-    if (counter != 0)               // right unbalanced
-    {
-        return false;
-    }
-    else                            // correct dotbracket
-    {
-        return true;
-    }
-}
-
-
-std::string only_paired(std::string dot_bracket)
-{
-    // removes the "." characters from the dot_bracket
-    assert(is_valid_dot_bracket(dot_bracket));  // only apply on legal dot_bracket
-    std::string ret = std::string(dot_bracket);
-    ret.erase(std::remove(ret.begin(), ret.end(), '.'), ret.end());
-    return ret;
-}
-
-
-int levenshtein(std::string a, std::string b)
-{
-    // string edit distance
-    int aLength = a.size();
-    int bLength = b.size();
-
-    std::vector< std::vector<int> > levDistances = std::vector< std::vector<int> >();
-    for(int i = 0; i < aLength+1; ++i)
-    {
-        levDistances.push_back(std::vector<int>(bLength+1, 0));
-    }
-
-
-    // fill first column
-    for (int i = 0; i < aLength+1; ++i)
-    {
-        levDistances[i][0] = i;
-    }
-
-    // fill second column
-    for (int j = 0; j < bLength+1; ++j)
-    {
-        levDistances[0][j] = j;
-    }
-
-    // fill the table now
-    for (int i = 0; i < aLength; ++i)
-    {
-        for (int j = 0; j < bLength; ++j)
-        {
-            if (a[i] == b[j])
-            {
-                levDistances[i + 1][j + 1] = levDistances[i][j];
-            }
-            else
-            {
-                int minDist = levDistances[i][j + 1];
-                int dist = levDistances[i + 1][j];
-                if (dist < minDist)
-                    minDist = dist;
-                dist = levDistances[i][j];
[…]

### 2015-06-22 src/RNAConsensus.cpp

@@ -1,5 +1,70 @@
 #include "../include/RNAConsensus.h"
 
+/***********************************************************************\
+ *
+ * File:           RngStream.cpp for multiple streams of Random Numbers
+ * Language:       C++ (ISO 1998)
+ * Copyright:      Pierre L'Ecuyer, University of Montreal
+ * Notice:         This code can be used freely for personal, academic,
+ *                 or non-commercial purposes. For commercial purposes, 
+ *                 please contact P. L'Ecuyer at: lecuyer@iro.umontreal.ca
+ * Date:           14 August 2001
+ *
+\***********************************************************************/
+
+int levenshtein(std::string a, std::string b)
+{
+    // string edit distance
+    int aLength = a.size();
+    int bLength = b.size();
+
+    std::vector< std::vector<int> > levDistances = std::vector< std::vector<int> >();
+    for(int i = 0; i < aLength+1; ++i)
+    {
+        levDistances.push_back(std::vector<int>(bLength+1, 0));
+    }
+
+
+    // fill first column
+    for (int i = 0; i < aLength+1; ++i)
+    {
+        levDistances[i][0] = i;
+    }
+
+    // fill second column
+    for (int j = 0; j < bLength+1; ++j)
+    {
+        levDistances[0][j] = j;
+    }
+
+    // fill the table now
+    for (int i = 0; i < aLength; ++i)
+    {
+        for (int j = 0; j < bLength; ++j)
+        {
+            if (a[i] == b[j])
+            {
+                levDistances[i + 1][j + 1] = levDistances[i][j];
+            }
+            else
+            {
+                int minDist = levDistances[i][j + 1];
+                int dist = levDistances[i + 1][j];
+                if (dist < minDist)
+                    minDist = dist;
+                dist = levDistances[i][j];
+                if (dist < minDist)
+                    minDist = dist;
+
+                levDistances[i + 1][j + 1] = minDist + 1;
+            }
+        }
+    }
+    return levDistances[aLength][bLength];
+}
+
+
+
 
 
 int get_subopt_length(std::string line)

### 2015-06-22 src/Solver.cpp

@@ -1,6 +1,12 @@
 #include "../include/Solver.h"
 
 
+bool Solver::is_silent()
+{
+    return silent_;
+}
+
+
 std::vector<int> select_random_genes(std::vector<Range> ranges, RngStream *prng)
 {
     // randomly choose a gene from within each set of genes (represented by intervals)

### 2015-06-22 src/SolverExact.cpp

@@ -6,9 +6,7 @@
 // constructors and destructors
 SolverExact::SolverExact(bool silent) : Solver(silent) { }
 
-
 SolverExact::~SolverExact() { }
-
 
 
 double find_smallest_distance_range(int self_index,
@@ -58,16 +56,17 @@ std::vector< std::vector<double> > elem_range_distance_matrix(std::vector<RangeI
 }
 
 
-bool is_leaf(SolutionNode sol)
+bool is_leaf(SubSolution sol)
 {
+    // checks wether or not a solution is complete
     return (sol.get_ranges().size() == 0);
 }
 
 
-void expand_solution(SolutionNode solution,
+void expand_solution(SubSolution solution,
                      std::vector< std::vector<double> > elem_range_dm,
                      std::vector< std::vector<double> > distance_matrix,
-                     std::priority_queue< SolutionNode > & search_space,
+                     std::priority_queue< SubSolution > & search_space,
                      double cutoff)
 {
     // used to create all children of a node
@@ -96,7 +95,7 @@ void expand_solution(SolutionNode solution,
         new_genes.push_back(new_gene);
 
         // create the new solution, score it and add it if its worth it
-        SolutionNode new_solution = SolutionNode(new_genes, new_ranges);
+        SubSolution new_solution = SubSolution(new_genes, new_ranges);
         new_solution.set_score(elem_range_dm, distance_matrix);
 
         if (new_solution.get_score() <= cutoff)
@@ -115,14 +114,12 @@ std::vector<Solution> SolverExact::solve(std::vector< std::vector<double> > dist
                                          std::vector<Range> ranges,
                                          unsigned long seeds[6]) const
 {
-    // could explore solutions by order of lowest
-    // possible cost (most promising) or otherwise
-    // only difference is the size the search space will get to
-    std::priority_queue<SolutionNode> search_space = std::priority_queue<SolutionNode>();
-    std::vector<SolutionNode> satisfying_solutions = std::vector<SolutionNode>();
+    // use a priority queue to represent the current search space
+    std::priority_queue<SubSolution> search_space = std::priority_queue<SubSolution>();
+    std::vector<SubSolution> satisfying_solutions = std::vector<SubSolution>();
     double best_score = std::numeric_limits<double>::max();
 
-    // sort the ranges by their size
+    // sort the ranges by their size (to avoid big branching at the start)
[…]

### 2015-06-22 src/SubSolution.cpp

@@ -1,11 +1,11 @@
 
-#include "../include/SolutionNode.h"
+#include "../include/SubSolution.h"
 
 
 // constructor and destructor
-SolutionNode::SolutionNode(std::vector<int> genes,
-                           std::vector<RangeIndex> ranges,
-                           double score)
+SubSolution::SubSolution(std::vector<int> genes,
+                         std::vector<RangeIndex> ranges,
+                         double score)
 {
     genes_ = genes;
     ranges_ = ranges;
@@ -13,7 +13,7 @@ SolutionNode::SolutionNode(std::vector<int> genes,
     return;
 }
 
-SolutionNode::SolutionNode(const SolutionNode& other)
+SubSolution::SubSolution(const SubSolution& other)
 {
     // copy constructor
     genes_ = other.get_genes();
@@ -22,39 +22,34 @@ SolutionNode::SolutionNode(const SolutionNode& other)
 }
 
 
-SolutionNode::~SolutionNode(){}
+SubSolution::~SubSolution(){}
 
 
 
 // getters and setters
-const std::vector<int> & SolutionNode::get_genes() const
+const std::vector<int> & SubSolution::get_genes() const
 {
     return genes_;
 }
 
-void SolutionNode::set_gene(int position, int new_gene)
+void SubSolution::set_gene(int position, int new_gene)
 {
     genes_[position] = new_gene;
     return;
 }
 
-void SolutionNode::add_gene(int new_gene)
-{
-    genes_.push_back(new_gene);
-    return;
-}
 
-std::vector<RangeIndex> SolutionNode::get_ranges() const
+std::vector<RangeIndex> SubSolution::get_ranges() const
 {
     return ranges_;
 }
 
-double SolutionNode::get_score() const
+double SubSolution::get_score() const
 {
     return score_;
 }
 
-void SolutionNode::set_score(std::vector< std::vector<double> > elem_range_dm,
+void SubSolution::set_score(std::vector< std::vector<double> > elem_range_dm,
                              std::vector< std::vector<double> > distance_matrix)
 {
     // set the score to the best possible score for this solution
@@ -89,27 +84,27 @@ void SolutionNode::set_score(std::vector< std::vector<double> > elem_range_dm,
 }
 
 
-Solution SolutionNode::return_solution()
+Solution SubSolution::return_solution()
 {
-    // convert to a solution and return it
+    // convert a subsolution to a solution and return it
[…]

### 2015-06-22 src/Test.cpp

@@ -1,129 +1,0 @@
-#include "Test.h"
-
-Test::Test()
-{
-    //ctor
-}
-
-Test::~Test()
-{
-    //dtor
-}
-
-
-bool test_RNAConsensus() {
-    bool working = true;
-    // step 1: reading from file
-    vector<vector<string> > data = read_data("../../../data/others/DM1-33.txt");
-    for(int i =0 ; i != data.size(); ++i) {
-        std::cout << std::endl;
-        for(int j= 0; j < data[i].size(); ++j) {
-            std::cout << data[i][j] << std::endl;
-        }
-    }
-    if (data.size() == 0)
-    {
-        working = false;
-    }
-    // step 2: converting dotbrackets to trees in efficient fashion
-    vector< vector<Tree*> > trees = get_tree_lists(data);
-    if (trees.size() == 0)
-    {
-        working = false;
-    }
-
-    // step 3: filtering
-    vector<string> consensus = vector<string>();
-    for(unsigned int i = 0; i < trees.size(); ++i)
-    {
-        consensus.push_back(trees[i][0]->get_brackets());
-    }
-    vector< vector<string> > filtered = filter_dot_brackets(data, consensus);
-
-
-    for(unsigned int i = 0; i < consensus.size(); ++i)
-    {
-        for(unsigned int j = 0; j < filtered[i].size(); ++j)
-        {
-            if ( consensus[i] != only_paired(filtered[i][j]) )
-            {
-                working = false;
-                break;
-            }
-        }
-    }
-
-    // free structures
-    for(size_t i = 0; i < trees.size(); ++i)
-    {
-        for (size_t j = 0; j < trees[i].size(); ++j)
-        {
-            delete (trees[i][j]);
-        }
-    }
-    trees.clear();
-    return working;
-}
-
-
-bool test_ConsensusProblem() {
-
-    // flag
-    bool working = true;
-
-    // fetch the data
-    vector<vector<string> > data = read_data("../../../data/others/DM1-33.txt");
-    vector<vector<Tree*> > trees = get_tree_lists(data);
-
-    // create the problem instance
-    ConsensusProblem<Tree*>* prob1 = new ConsensusProblem<Tree*>(trees, *unit_distance);
-    ConsensusProblem<string>* prob2 = new ConsensusProblem<string>(data, *levenshtein);
-
-    vector<vector<double> > mat = prob2->get_distance_matrix();
[…]

### 2015-06-22 src/Tree.cpp

@@ -9,6 +9,79 @@
 using std::vector;
 
 
+bool is_valid_dot_bracket(std::string dot_bracket)
+{
+    // tests Vienna dot-bracket for illegal structure (or symbol)
+    int counter = 0;
+
+    for (unsigned int i = 0; i < dot_bracket.size(); ++i)
+    {
+        char c = dot_bracket[i];
+        if (c == '(')               // stack
+        {
+            counter += 1;
+        }
+        else if (c == ')')          // unstack
+        {
+            counter -= 1;
+        }
+        else if (c != '.')          // illegal character
+        {
+            return false;
+        }
+        if (counter < 0)            // left unbalanced
+        {
+            return false;
+        }
+    }
+    if (counter != 0)               // right unbalanced
+    {
+        return false;
+    }
+    else                            // correct dotbracket
+    {
+        return true;
+    }
+}
+
+
+std::string only_paired(std::string dot_bracket)
+{
+    // removes the "." characters from the dot_bracket
+    assert(is_valid_dot_bracket(dot_bracket));  // only apply on legal dot_bracket
+    std::string ret = std::string(dot_bracket);
+    ret.erase(std::remove(ret.begin(), ret.end(), '.'), ret.end());
+    return ret;
+}
+
+Node* dot_bracket_to_node(std::string dot_bracket)
+{
+    //creates a base pair tree from Vienna dot bracket
+    assert(is_valid_dot_bracket(dot_bracket));
+    Node* root = new Node();
+    Node* position = root;
+    int node_id = 0;
+    for (int i = 0; dot_bracket[i] != '\0'; i++)
+    {
+        char c = dot_bracket[i];
+        if (c == '(')
+        {
+            Node* child = new Node(position, node_id);
+            node_id = node_id + 1;
+            position = child;
+        }
+        else if (c == ')')
+        {
+            position = position->get_parent();
+        }
+        else
+        {
+            continue;
+        }
+    }
+    return root;
+}
+
 Tree::Tree(std::string brackets)
 {
     // all trees can be represented by a bracket (Vienna dot bracket minus the dots)

### 2015-06-22 include/ConsensusProblem.h
removed the Monte Carlo solver, now simpler

@@ -31,9 +31,9 @@ public:
 
 
 private:
-    std::vector<T> objects;
-    std::vector<Range> ranges;
-    std::vector< std::vector<double> > distance_matrix;
+    std::vector<T> objects_;
+    std::vector<Range> ranges_;
+    std::vector< std::vector<double> > distance_matrix_;
 };
 
 // used to assert that the distance matrix passed

### 2015-06-22 src/ConsensusProblem.cpp

@@ -9,26 +9,26 @@ ConsensusProblem<T>::ConsensusProblem(std::vector< std::vector<T> > data,
                                       int (*distance_function)(T a, T b))
 {
     // constructor
-    objects = vector<T>();
-    ranges = vector<Range>();
+    objects_ = vector<T>();
+    ranges_ = vector<Range>();
     int cumul = 0;
 
     for (int x = 0; x < data.size(); ++x)
     {
         for (int y = 0; y < data[x].size(); ++y)
         {
-            objects.push_back(data[x][y]);
+            objects_.push_back(data[x][y]);
         }
         Range r = {cumul, (int) (cumul + data[x].size())};
-        ranges.push_back( r );
+        ranges_.push_back( r );
         cumul += data[x].size();
     }
 
     // fill the distance matrix
-    this->distance_matrix = vector< vector<double> >();
-    for(int i = 0; i < objects.size(); ++i)
+    distance_matrix_ = vector< vector<double> >();
+    for(int i = 0; i < objects_.size(); ++i)
     {
-        this->distance_matrix.push_back(vector<double>(objects.size(), std::numeric_limits<double>::infinity()));
+        distance_matrix_.push_back(vector<double>(objects_.size(), std::numeric_limits<double>::infinity()));
     }
 
     // memoize distance calculations
@@ -36,24 +36,24 @@ ConsensusProblem<T>::ConsensusProblem(std::vector< std::vector<T> > data,
     // function must be metric!
     // create dict and prefill it
     map < T, map <T, double> > memoized_distances = map < T, map <T, double> >();
-    for (int i = 0; i < objects.size(); ++i)
+    for (int i = 0; i < objects_.size(); ++i)
     {
-        memoized_distances[objects[i]] = map<T, double>();
+        memoized_distances[objects_[i]] = map<T, double>();
 
     }
 
     // only compare those not of the same group
     // compute distances only once for each pair of objects
     // to do so, objects must be hashable and have equality operators (in this implementation at least)
-    for(int i = 0; i < ranges.size()-1; ++i)
+    for(int i = 0; i < ranges_.size()-1; ++i)
     {
-        int low1 = ranges[i].low;
-        int high1 = ranges[i].high;
+        int low1 = ranges_[i].low;
+        int high1 = ranges_[i].high;
 
-        for(int j = i+1; j < ranges.size(); ++j)
+        for(int j = i+1; j < ranges_.size(); ++j)
         {
[…]

### 2015-06-22 src/MCCONS_MC.cpp

@@ -1,48 +1,0 @@
-#include "../include/MCCONS.h"
-#include "../include/OptionParser.h"
-#include "../include/SolverMC.h"
-
-using optparse::OptionParser;
-
-
-int main(int argc, char *argv[])
-{
-    OptionParser parser = OptionParser().description("MC-Cons Consensus Optimizer Using a Monte Carlo and Steepest Descent");
-    parser.add_option("-f", "--data").dest("data_file").help("path to MARNA-like input file");
-    parser.add_option("-n", "--num_trials").dest("num_trials").help("number of trials").type("size_t");
-    parser.add_option("-s", "--silent").action("store_true").dest("silent").help("don't display status to stderr");
-    optparse::Values options = parser.parse_args(argc, argv);
-
-
-    // GLOBAL SETTINGS (MODIFY AT YOUR OWN RISKS (WHICH ARE MINIMAL))
-    unsigned long SEEDS[6] = {42, 42, 42, 42, 42, 42};
-    size_t NUM_TRIALS = 500;
-
-    // extract parameters and lauch the genetic algorithm (either silent or not)
-    if (options.is_set("data_file"))
-    {
-      // extract command line parameters
-      bool silent = false;
-      if (options["silent"] == "1")
-      {
-          silent = true;
-      }
-      std::string path = options["data_file"];
-      if (options.is_set("num_trials"))
-      {
-          NUM_TRIALS = atoi(options["num_trials"].c_str());
-      }
-
-      // instantiate the genetic algorithm solver
-      Solver* solver = new SolverMC(NUM_TRIALS, silent);
-
-      // execute and cleanup
-      MCCONS(path, solver, SEEDS);
-      delete solver;
-    } else
-    {
-      // something went wrong with the arguemnts, print error message and exit
-      std::cerr << "Error: something went wrong, please check usage -h or --help" << std::endl;
-      std::exit(0);
-    }
-}

### 2015-07-15 slides/presentation/presentation.tex
did major work on the presentation

\documentclass{beamer}
\DeclareGraphicsExtensions{.pdf,.png,.jpg,.eps}
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\include{graphics}

\title{MC-Cons: RNA 2D structural assignment}
\author{Gabriel C-Parent}
%\date{project name}

\usetheme{Execushares}

\begin{document}

\maketitle

\begin{frame}
	\frametitle{outline}
	\begin{center}
		\begin{itemize}
			\item what?
			\item why?
			\item how?
			\item what's next?
			\item Q\&A
		\end{itemize}
	\end{center}
\end{frame}

\section{what?}

\subsection{graphical example}

\begin{frame}
	\frametitle{so you have structures...}
\begin{figure}[!htb]
\centering
\resizebox{0.5\textwidth}{!}{\input{all_subopts.pdf_tex}}
\caption{Given secondary structures of functionally related RNAs, which ones might explain the function?}
\end{figure} 
\end{frame}

\begin{frame}
	\frametitle{now you can have consensus!}
	\begin{figure}[!htb]
	\centering
	\resizebox{0.75\textwidth}{!}{\input{consensus_example.pdf_tex}}
	\caption{Let's suppose that their function is explained by a subset of similar structures (consensus). \\\hspace{\textwidth}In this case, 2 consensus make sense. Sometimes you get more, sometimes you get less.}
	\end{figure} 
\end{frame}

\section{why?}

\begin{frame}
	\frametitle{what's the point?}
\begin{figure}[!htb]
\centering
\resizebox{0.55\textwidth}{!}{\input{consensus_banner.pdf_tex}}
\end{figure}

	\begin{itemize}

		\item goal: finding common/similar structures within suboptimal secondary structures
		\item useful for
		\begin{itemize}
			\item improving secondary structure prediction for functionally related RNAs
			\item exploring shared alternative structures
		\end{itemize}
	\end{itemize}

\end{frame}

\section{how?}

\subsection{general workflow}
\begin{frame}
	\frametitle{the workflow}
	\begin{columns}
		\column{0.38\linewidth}
		\centering
		\begin{figure}[!htb]
			\resizebox{1.\textwidth}{!}{\input{workflow.pdf_tex}}
		\end{figure}

        \column{0.58\linewidth}
        \begin{itemize}
        		\item a consensus minimizes the sum of pairs of distances between objects
        		\item optimization steps
        		\begin{enumerate}
        		        			\item base pair structural similarity (modified tree edit distance)
        			\item string edit distance on Vienna dot-bracket representation
        		\end{enumerate}

        \end{itemize}
         \end{columns}

\end{frame}

\subsection{matching tree base pairs}
\begin{frame}
	\frametitle{first, we match base pairs structure}
	\begin{figure}[!htb]
	\centering
	\resizebox{0.4\textwidth}{!}{\input{bp_tree.pdf_tex}}
	\end{figure}

	\begin{itemize}
		\item base pair trees created (many-to-one function)
		\item unit tree indel distance calculates number of base pairs to delete or create to transform a tree into another
	\end{itemize}

\end{frame}

\subsection{matching strings}
\begin{frame}
	\frametitle{refine by string edit distance}

\end{frame}

\section{examples}
\subsection{IREs}
\begin{frame}
	\frametitle{IREs}
	\begin{figure}[!htb]
	\centering
	\resizebox{0.75\textwidth}{!}{\input{ires_consensus.pdf_tex}}
	\caption{IREs structural assignment given 10 suboptimal for 29 different IRES.}
	\end{figure} 
\end{frame}

\subsection{tRNAs}
\begin{frame}
	\frametitle{tRNAs}

\end{frame}

\section{what's next?}
\begin{frame}
	\frametitle{future work}
	\begin{itemize}
		\item MC-Cons improvements
		\begin{itemize}
			\item parallelization
			\item better distance functions?
			\item heuristic improvement?
		\end{itemize}

		\item derivative works
		\begin{itemize}
			\item 3D to 2D tree distance
			\item base pair tree constraints/masks?
		\end{itemize}

	\end{itemize}
\end{frame}

\section{Conclusion}

\begin{frame}
	\frametitle{Conclusion}
	\begin{itemize}
		\item Adversarial IR
		\begin{itemize}
			\item it's hard
		\end{itemize}
		\item Link Analysis/Mining
		\begin{itemize}
			\item wide variety of problems and domains
		\end{itemize}
		\item Link Analysis for Web
		\begin{itemize}
			\item Centrality
			\item Pagerank
			\item Topic-sensitive PageRank
			\item Hubs and Authorities
		\end{itemize}
		\item Still many challenges
	\end{itemize}
\end{frame}

\section{Q\&A}

\end{document}

### 2015-07-15 slides/data/util.py

@@ -1,0 +1,274 @@
+"""interface with flashfold RNA secondary structure folding tool"""
+
+import copy
+import os
+import subprocess
+import shlex
+import math
+from xml.dom import minidom
+from collections import Counter
+import matplotlib.cm
+import matplotlib.colors
+
+
+VARNA_PATH = "VARNAv3-93.jar"
+
+def call_command(command, pipe=None, echo=False):
+    """simple shell call interface for python"""
+    if echo:
+        print command
+
+    process = subprocess.Popen(shlex.split(command.encode("ascii")),
+                               stdin=subprocess.PIPE,
+                               stdout=subprocess.PIPE,
+                               stderr=subprocess.PIPE)
+
+    output = process.communicate(input=pipe)
+    return output
+
+
+def read_marna(file_name):
+    """read marna-like file format"""
+    data = []
+    with open(file_name) as data_file:
+        name = ""
+        seq = ""
+        subopts = []
+        lines = data_file.readlines()
+        for line in lines:
+            if line.startswith(">"):  # new name
+                if name != "":
+                    data.append((name, seq, subopts))
+                name = line.strip()[1:]  # remove ">" char
+                seq = ""
+                subopts = []
+            stripped = line.strip()[0]
+            if stripped in "AUGC":  # new sequence
+                seq = line.strip()
+            elif stripped in ".()":  # new subopt
+                splitted = line.strip().split()
+                subopt, _ = splitted[0], float(splitted[1])
+                subopts.append(subopt)
+            else:
+                continue
+        if name != "":  # end of file
+            data.append((name, seq, subopts))
+    return data
+
+
+def preprocess_data(data, limit_subopts=float("inf")):
+    """applies some modifs to data, making it easier to work with VARNA"""
+    data = copy.copy(data)  # eliminate side effects
+
+    # limit the number of subopts if wished
+    if limit_subopts < float("inf"):
+      data = [(i[0], i[1], i[2][:limit_subopts]) for i in data]
+
+    # put it in list format (easier to call varna with this)
+    data_list = []
+    for name, seq, subopts in data:
+      tmp = []
+      for i, subopt in enumerate(subopts):
+        tmp.append((name+"_{}".format(i), seq, subopt))
+      data_list.append(tmp)
[…]

### 2015-07-15 slides/util.py

@@ -1,160 +1,0 @@
-"""interface with flashfold RNA secondary structure folding tool"""
-
-import copy
-import subprocess
-import shlex
-import xml.etree
-import math
-
-VARNA_PATH = "../../VARNAv3-93.jar"
-
-def call_command(command, pipe=None, echo=False):
-    """simple shell call interface for python"""
-    if echo:
-        print command
-
-    process = subprocess.Popen(shlex.split(command.encode("ascii")),
-                               stdin=subprocess.PIPE,
-                               stdout=subprocess.PIPE,
-                               stderr=subprocess.PIPE)
-
-    output = process.communicate(input=pipe)
-    return output
-
-
-def read_data(file_name):
-    """read marna-like file format"""
-    data = []
-    with open(file_name) as data_file:
-        name = ""
-        seq = ""
-        subopts = []
-        lines = data_file.readlines()
-        for line in lines:
-            if line.startswith(">"):  # new name
-                if name != "":
-                    data.append((name, seq, subopts))
-                name = line.strip()[1:]  # remove ">" char
-                seq = ""
-                subopts = []
-            stripped = line.strip()[0]
-            if stripped in "AUGC":  # new sequence
-                seq = line.strip()
-            elif stripped in ".()":  # new subopt
-                splitted = line.strip().split()
-                subopt, _ = splitted[0], float(splitted[1])
-                subopts.append(subopt)
-            else:
-                continue
-        if name != "":  # end of file
-            data.append((name, seq, subopts))
-    return data
-
-
-def preprocess_data(data, limit_subopts=float("inf")):
-    """applies some modifs to data, making it easier to work with VARNA"""
-    data = copy.copy(data)  # eliminate side effects
-
-    # limit the number of subopts if wished
-    if limit_subopts < float("inf"):
-      data = [(i[0], i[1], i[2][:limit_subopts]) for i in data]
-
-    # put it in list format (easier to call varna with this)
-    data_list = []
-    for name, seq, subopts in data:
-      tmp = []
-      for i, subopt in enumerate(subopts):
-        tmp.append((name+"_{}".format(i), seq, subopt))
-      data_list.append(tmp)
-    return data_list
-
-
-def call_VARNA(name, sequence, structure, bpcolor, options=""):
[…]

### 2015-07-15 src/cluster.py

@@ -350,7 +350,7 @@ def clustermap(data, pivot_kws=None, method='average', metric='euclidean',
 #show_solution(tree_solutions[0], tree_problem, cluster_method="complete")
 
 # look at the first dot bracket solution
-ex = show_solution(dot_bracket_solutions[0][0], dot_bracket_problems[0], cluster_method="complete")
+#ex = show_solution(dot_bracket_solutions[0][0], dot_bracket_problems[0], cluster_method="complete")
 
 
-ex.savefig("../../article/figs/consensus.eps")
+#ex.savefig("../../article/figs/consensus.eps")

### 2015-07-18 slides/presentation/presentation.tex
working on aknowledgments

@@
 \begin{frame}
-	\frametitle{the workflow}
+	\frametitle{workflow}
 	\begin{columns}
@@
         		\end{enumerate}
-
+			\item two solver in C++ (branch \& bound and genetic algorithm)
         \end{itemize}
@@
 
-\subsection{matching tree base pairs}
+\subsection{matching base pairs}
 \begin{frame}
-	\frametitle{first, we match base pairs structure}
+	\frametitle{first match base pairs}
 	\begin{figure}[!htb]
@@
 		\item base pair trees created (many-to-one function)
-		\item unit tree indel distance calculates number of base pairs to delete or create to transform a tree into another
+		\item unit tree indel distance calculates number of base pairs to add or remove to transform a tree into another
 	\end{itemize}
@@
 
-\subsection{matching strings}
+\subsection{match the bulges}
 \begin{frame}
-	\frametitle{refine by string edit distance}
-
+	\frametitle{then match unpaired}
+	\begin{figure}[!htb]
+	\centering
+	\resizebox{0.75\textwidth}{!}{\input{tree_filtering.pdf_tex}}
+	\end{figure}	
+	\begin{itemize}
+		\item filtered by tree structure, now refine
+		\item placing the bulges around the skeleton...
+		\item string edit distance on dot-bracket
+	\end{itemize}
 \end{frame}
@@
 		\begin{itemize}
+			\item base pair tree constraints/masks
 			\item 3D to 2D tree distance
-			\item base pair tree constraints/masks?
 		\end{itemize}
@@
 
-\section{Conclusion}
+\section{recap}
 
 \begin{frame}
-	\frametitle{Conclusion}
+	\frametitle{recap}
 	\begin{itemize}
-		\item Adversarial IR
-		\begin{itemize}
-			\item it's hard
-		\end{itemize}
-		\item Link Analysis/Mining
-		\begin{itemize}
-			\item wide variety of problems and domains
-		\end{itemize}
-		\item Link Analysis for Web
-		\begin{itemize}
-			\item Centrality
-			\item Pagerank
-			\item Topic-sensitive PageRank
-			\item Hubs and Authorities
-		\end{itemize}
-		\item Still many challenges
+		\item MC-Cons creates RNA 2D structural assignment
+		\item useful for finding structures explaining function
+		\item available at \url{https://github.com/major-lab/MC-Cons}
 	\end{itemize}
@@
 
+\begin{frame}
+	\frametitle{Q\&A}
+		\url{https://github.com/major-lab/MC-Cons}
+\end{frame}
+
 \end{document}

### 2015-08-03 doc/article/article.tex
changed directory structure

--- 2015-06-21 article/article.tex
@@
 \section*{Introduction}
-Lorem ipsum dolor %sit~\cite{bib1}%%
-amet, consectetur adipiscing elit.
+
+% context
+
+%Lorem ipsum dolor %sit~\cite{bib1}%%
+%amet, consectetur adipiscing elit.
 Curabitur eget porta erat.
@@
 
-\begin{equation}\label{eq:schemeP} 
-D_{coll} = \frac{D_f+\frac{[S]^2}{K_D S_T} D_S} {1+\frac{[S]^2}{K_D S_T}}, 
-D_{sm} = \frac{D_f+ \frac{[S]}{K_D} D_S}{1+\frac{[S]}{K_D}},
-\end{equation}
+%\begin{equation}\label{eq:schemeP} 
+%D_{coll} = \frac{D_f+\frac{[S]^2}{K_D S_T} D_S} {1+\frac{[S]^2}{K_D S_T}}, 
+%D_{sm} = \frac{D_f+ \frac{[S]}{K_D} D_S}{1+\frac{[S]}{K_D}},
+%\end{equation}
 
@@
 
-\begin{enumerate}
-\item{react}
-\item{diffuse free particles}
-\item{increment time by dt and go to 1}
-\end{enumerate}
+%\begin{enumerate}
+%\item{react}
+%\item{diffuse free particles}
+%\item{increment time by dt and go to 1}
+%\end{enumerate}
 
 % Results and Discussion can be combined.
+\newpage
 \section*{Results}
@@
 
+\newpage
 \section*{Discussion}
@@
 It is distributed at GitHub, https://github.com/major-lab/MC-Cons along with documentation and working examples.
-The algorithms are implemented in C++ with no external dependencies and offers a very simple command line interface.
+The algorithms are implemented in portable ISO 1998 C++ without external dependencies and offers a very simple command line interface.
 
+\newpage
 \section*{Supporting Information}

### 2015-08-06 doc/article/article.tex
linted the util.py and did some work on clustering

@@
 {\Large
-\textbf\newline{MC-Cons: RNA secondary structure assignments}
+\textbf\newline{MC-Cons: RNA secondary structural assignments}
 }
@@
 
+\subsection*{The Workflow}
+
+% within tree consensus, minimizing a naive distance: string edit distance
+MC-Cons is a two-step optimization procedure (Fig.~\ref{fig1}).
+
+First, \textit{base pair tree consensus} are calculated.
+This insures that the final consensus will minimize the difference in base pair topology.
+For each base pair tree consensus and for each group of structures, the subset of secondary structures whose base pair tree fits the one assigned to that group is kept.
+
+String edit distance consensus (on Vienna dot brackets) are then calculated.
+Normally this distance function would not be sufficient, but since tree distance is already minimized, the final consensus is satisfactory.\newline
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[scale=0.4]{figs/mccons_flowchart2.eps}
+\caption{{\bf Flowchart illustration of MC-Cons procedure.}
+The two-step procedure is quite simple. The user only inputs data and chooses which consensus solver, either the exact one for small instances or the genetic algorithm for larger problems.}
+\label{fig1}
+\end{center}
+\end{figure}
+
 \subsection*{Base Pair Tree Representation of Secondary Structure}
@@
 Although this representation greatly increases the size of the auxiliary space (codomain) compared to level 5 abstract shapes, it does not suffer from major changes in representation if small stem-loops are added or removed.
-It also does not map structures of widely varying stem length to the same element in the codomain.
-
-%example of the conversion process 
-\noindent The mapping from a Vienna dot bracket representation can be made simply by removing the dots (unpaired nucleotides) and then creating a tree from it.
-Since all nodes represent base pairs, the resulting tree can be considered unlabelled.
-Note that an artificial root is added to all structures.
 
@@
 % relation to the tree edit distance (special scoring scheme)
+A tree edit distance function is used to compare base pair trees.
+
+Insertions and deletions of base pairs have unit cost and since all nodes have the same label there is no relabelling.
+
+This distance function is called the unit tree indel distance It is referred to as a tree indel distance and 
 The unit tree indel distance is a simple and intuitive distance function based on the well known tree edit distance \cite{tai, zhang_shasha}, where insertions and deletions have a unit cost.
@@
 
-Minimizing the sum of pairwise distance between sets of objects is far from new and a significant body of work has been done in the field of location theory \cite{compact_location}. \newline
+Minimizing the sum of pairwise distance between sets of objects has been extensively investigated in the field of location theory \cite{compact_location}. Indeed, each step of the \newline
 
@@
 The operator applies a small number of iterations of local search with a certain probability.
-
-\subsection*{MC-Cons}
-
-% within tree consensus, minimizing a naive distance: string edit distance
-MC-Cons is a two-step optimization procedure (Fig.~\ref{fig1}).
-First, \textit{base pair tree consensus} are calculated.
-This insures that the final consensus will minimize the difference in base pair topology.
-For each base pair tree consensus and for each group of structures, the subset of secondary structures whose base pair tree fits the one assigned to that group is kept.
-String edit distance consensus (on Vienna dot brackets) are then calculated.
-Normally this distance function would not be sufficient, but since tree distance is already minimized, the final consensus is satisfactory.\newline
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[scale=0.4]{figs/mccons_flowchart2.eps}
-\caption{{\bf Flowchart illustration of MC-Cons procedure.}
-The two-step procedure is quite simple. The user only inputs data and chooses which consensus solver, either the exact one for small instances or the genetic algorithm for larger problems.}
-\label{fig1}
-\end{center}
-\end{figure}
 

### 2015-08-06 src/cluster.py

@@ -1,3 +1,6 @@
+#! /usr/bin/env python
+"""Clustering functions used when pretty printing consensus"""
+
 # some clustering utilities
 import scipy.cluster
 import seaborn

### 2015-08-06 src/minimal_spanning_tree.py

@@ -1,0 +1,103 @@
+"""applying Prim's algorithm on a distance matrix (pure python)"""
+
+
+import copy
+
+
+class LabeledDistanceMatrix(object):
+    """keeps labels, distance matrix, and a min heap"""
+    def __init__(self, distance_matrix, labels=None):
+        # check that the matrix is symmetric
+        if not distance_matrix_is_symmetric(distance_matrix):
+            raise Exception("Non symmetric distance matrix")
+
+        # exclude self connections
+        self.distance_matrix = copy.copy(distance_matrix)
+        for i in range(len(self.distance_matrix)):
+            self.distance_matrix[i][i] = float('inf')
+        self.labels = labels
+
+
+    def __str__(self):
+        """ display the distance matrix and its fields"""
+        to_str = "Distance Matrix\n"
+        if not self.labels is None:
+            to_str += "Labels : {}\n".format(self.labels)
+            to_str += "\n"
+        for i in self.distance_matrix:
+            to_str += str(i)
+            to_str += "\n"
+        return to_str
+
+
+    def __repr__(self):
+        return str(self)
+
+
+    def update_distance_matrix(self, row):
+        """set all the distance on the row to infinity"""
+        for i in range(len(self.distance_matrix)):
+            self.distance_matrix[row][i] = float('inf')
+        return
+
+
+    def find_min_arc(self, selected_columns):
+        """search for best feasible arc and update the matrix"""
+        best_column = -1
+        best_row = -1
+        best_distance = float('inf')
+        for column in selected_columns:
+            for row in range(len(self.distance_matrix)):
+                current_dist = self.distance_matrix[row][column]
+                if current_dist < best_distance:
+                    best_distance = current_dist
+                    best_column = column
+                    best_row = row
+        # kill the row
+        self.update_distance_matrix(best_row)
+        return (best_column, best_row)
+
+
+def distance_matrix_is_symmetric(distance_matrix):
+    """checks wether or not the distance matrix is symmetric"""
+    is_symmetric = True
+    for i in range(len(distance_matrix)):
+        if is_symmetric:
+            for j in range(len(distance_matrix)):
+                if distance_matrix[i][j] != distance_matrix[j][i]:
+                    is_symmetric = False
[…]

### 2015-08-06 src/shape_explorer.py

@@ -3,220 +3,7 @@
 
 
 import argparse
-
-
-def is_valid_dot_bracket(dot_bracket):
-    """tests Vienna dot-bracket for illegal structure (or symbol)"""
-    counter = 0
-    for i in dot_bracket:
-        if i == '(':
-            counter += 1
-        elif i == ')':
-            counter -= 1
-        elif i != '.':  # illegal symbol
-            return False
-        if counter < 0:  # unbalanced structure
-            return False
-    if counter != 0:
-        return False  # unbalanced structure
-    return True
-
-
-class Consensus(object):
-    """simple object to hold all the information about a consensus"""
-    def __init__(self, number, tree_dist, bp_dist, subopts, level):
-        self.number = number
-        self.tree_dist = tree_dist
-        self.bp_dist = bp_dist
-        for subopt in subopts:
-            assert is_valid_dot_bracket(subopt), "problematic structure: {}".format(subopt)
-        self.subopts = subopts
-        self.abstract_shapes = ([dot_bracket_to_abstract_shape(subopt, level)
-                                for subopt in self.subopts])
-
-    def __str__(self):
-        ret = "> {0} {1} {2}\n".format(self.number,
-                                       self.tree_dist,
-                                       self.bp_dist)
-        for subopt in self.subopts:
-            ret += "{0}\n".format(subopt)
-        ret += "\n"
-        return ret
-
-    def __repr__(self):
-        return str(self)
-
-    def str_with_shapes(self):
-        """string representation with abstract shapes along suboptimals"""
-        ret = "> {0} {1} {2}\n".format(self.number,
-                                self.tree_dist,
-                                self.bp_dist)
-        for subopt, shape in zip(self.subopts, self.abstract_shapes):
-            ret += "{0} {1}\n".format(subopt, shape)
-        ret += "\n"
-        return ret
-
-    def get_shape_signature(self):
-        """return string of shapes separated by newline"""
-        ret = ""
-        for shape in self.abstract_shapes:
-            ret += "{}\n".format(shape)
-        return ret
-
-    def __lt__(self, other):
-        if self.tree_dist == other.tree_dist:
-            return self.bp_dist < other.bp_dist
-        else:
-            return self.tree_dist < other.tree_dist
-
-    def __gt__(self, other):
-        return other < self
-
-    def __eq__(self, other): # equality on score
-        return ((self.tree_dist == other.tree_dist) and
-                (self.bp_dist == other.bp_dist))
-
-
[…]

### 2015-08-06 src/util.py

@@ -1,0 +1,433 @@
+"""Objects and functions used in multiple scripts"""
+
+
+
+import collections
+
+
+def is_valid_dot_bracket(dot_bracket):
+    """tests Vienna dot-bracket for illegal structure (or symbol)"""
+    counter = 0
+    for i in dot_bracket:
+        if i == '(':
+            counter += 1
+        elif i == ')':
+            counter -= 1
+        elif i != '.':  # illegal symbol
+            return False
+        if counter < 0:  # unbalanced structure
+            return False
+    if counter != 0:
+        return False  # unbalanced structure
+    return True
+
+
+
+def dot_bracket_to_abstract_shape(structure, level):
+    """Converts a Vienna dot-bracket structure into
+       its corresponding abstract shape as defined in RNAshapes.
+       Written by Stephen Leong Koan, IRIC, 2013"""
+    # the 3 levels we use are the following
+    # 1: Most accurate - all loops and all unpaired
+    # 3: Nesting pattern for all loop types but no unpaired regions
+    # 5: Most abstract - helix nesting pattern and no unpaired regions
+    assert level in [1, 3, 5]
+    list_opener = []
+
+    list_stems = []
+    i = 0
+    list_stem_start = []
+
+    # separate into stems
+    while i < len(structure):
+        if structure[i] == "(":
+            list_opener.append(i)
+            if not list_stem_start:
+                list_stem_start.append(i)
+        elif structure[i] == ")":
+            current_stem = dict(opener=[], closer=[], pairs=[])
+            while i < len(structure):
+                if structure[i] == ")":
+                    closer = list_opener.pop(-1)
+                    current_stem["opener"].append(closer)
+                    current_stem["closer"].append(i)
+                    current_stem["pairs"].append((closer, i))
+                    if closer in list_stem_start:
+                        break
+                elif structure[i] == "(":
+                    list_stem_start.append(i)
+                    i -= 1
+                    break
+                i += 1
+            list_stems.append(current_stem)
+        i += 1
+
+    # build the level 1 for each stems
+    range_occupied = []
+    dict_lvl1 = dict()
+    for stem in list_stems:
+        range_open = range(min(stem['opener']), max(stem['opener'])+1)
+        range_close = range(min(stem['closer']), max(stem['closer'])+1)
+        range_occupied.extend(range_open+range_close)
+
[…]

### 2015-08-10 src/shape_explorer.py
added clustering to output options

@@ -73,7 +73,7 @@ if __name__ == '__main__':
     # (with or without abstract shapes)
     STRFN = lambda x: x.__str__()
     if ARGS.display_abstract_shape == True:
-        STRFN = lambda x: x.str_with_shapes(ARGS.level)
+        STRFN = lambda x: x.str_with_shapes_order(ARGS.level)
 
     # output to either stdout or specified file
     if ARGS.output_path == "stdout":

### 2015-08-10 src/util.py

@@ -1,7 +1,6 @@
 """Objects and functions used in multiple scripts"""
 
 
-
 import collections
 
 
@@ -360,7 +359,7 @@ class Consensus(object):
     def __repr__(self):
         return str(self)
 
-    def str_with_shapes(self, level):
+    def str_with_shapes_order(self, level):
         """string representation with abstract shapes along suboptimals"""
         assert level in [5, 3, 1]
         level_to_index = [None, 2, None, 1, None, 0]
@@ -372,8 +371,11 @@ class Consensus(object):
         ret = "> {0} {1:.3f} {2:.3f}\n".format(self.number,
                                                avg_tree_dist,
                                                avg_string_dist)
-        for subopt, shape in zip(self.subopts, self.abstract_shapes):
-            ret += "{0} {1}\n".format(subopt, shape[access_index])
+        order = consensus_to_order(self)
+        for index in order:
+            subopt = self.subopts[index]
+            shape = self.abstract_shapes[index][access_index]
+            ret += "{0} {1} {2}\n".format(subopt, index, shape)
         ret += "\n"
         return ret
 
@@ -397,8 +399,10 @@ class Consensus(object):
             return (matrix_average(self.tree_dists) <
                     matrix_average(other.tree_dists))
 
+
     def __gt__(self, other):
         return other < self
+
 
     def __eq__(self, other): # equality on score
         equal_tree_dists = (matrix_average(self.tree_dists) ==
@@ -443,4 +447,184 @@ def read_consensus_file(file_name):
             consensus.append(Consensus(number,
                                        subopts))
     return consensus
+
+
+
+def _compute_weighted_distances(consensus,
+                                shape_weight=10000000.,
+                                tree_weight=100000.,
+                                string_weight=1.):
+    """converts the multiple distances used into a single distance
+       matrix by a simple weighted sum (many possibilities)"""
+
+    # first, we'd like to establish a distance between shapes
+    num_elems = len(consensus.subopts)
+    shape_dists = [[0. for _ in range(num_elems)] for _ in range(num_elems)]
+    # calculate the unit tree edit distance on the shapes
+    for index_1 in range(num_elems - 1):
+        (level_5_1, _, _) = consensus.abstract_shapes[index_1]
+        dot_bracket_1 = level_5_1.replace("[", "(")
+        dot_bracket_1 = dot_bracket_1.replace("]", ")")
+        tree_1 = dot_bracket_to_tree(dot_bracket_1)
[…]

### 2015-08-10 doc/article/article.tex
did some reformulation in the article and added few articles

@@
 \section*{Abstract}
-%Discovery and characterization of functional RNA structures remains challenging due to deficiencies in de novo secondary structure modeling.
-Here we describe a dynamic programming approach for model-free sequence comparison that incorporates high-throughput chemical probing data.
-Based on SHAPE probing data alone, ribosomal RNAs (rRNAs) from three diverse organisms – the eubacteria E.
-coli and C.
-difficile and the archeon H.
-volcanii – could be aligned with accuracies comparable to alignments based on actual sequence identity.
-When both base sequence identity and chemical probing reactivities were considered together, accuracies improved further.
-Derived sequence alignments and chemical probing data from protein-free RNAs were then used as pseudo-free energy constraints to model consensus secondary structures for the 16S and 23S rRNAs.
-There are critical differences between these experimentally-informed models and currently accepted models, including in the functionally important neck and decoding regions of the 16S rRNA.
-We infer that the 16S rRNA has evolved to undergo large-scale changes in base pairing as part of ribosome function.
-As high-quality RNA probing data become widely available, structurally-informed sequence alignment will become broadly useful for de novo motif and function discovery.
+
+We consider the problem of selecting 
 RNA secondary structural assignment is an interesting problem. Bla bla bla.
@@
 \newpage
-\section*{Methods}
+\section*{Design and Implementation}
 
-\subsection*{The Workflow}
+\subsection*{Consensus Problem}
 
-% within tree consensus, minimizing a naive distance: string edit distance
-MC-Cons is a two-step optimization procedure (Fig.~\ref{fig1}).
+Given $n$ sets of $ m_1, m_2... m_n$ objects, the consensus problem is to find a set of $n$ objects, one per set, minimizing the sum of their pairwise distance.
 
-First, \textit{base pair tree consensus} are calculated.
-This insures that the final consensus will minimize the difference in base pair topology.
-For each base pair tree consensus and for each group of structures, the subset of secondary structures whose base pair tree fits the one assigned to that group is kept.
-
-String edit distance consensus (on Vienna dot brackets) are then calculated.
-Normally this distance function would not be sufficient, but since tree distance is already minimized, the final consensus is satisfactory.\newline
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[scale=0.4]{figs/mccons_flowchart2.eps}
-\caption{{\bf Flowchart illustration of MC-Cons procedure.}
-The two-step procedure is quite simple. The user only inputs data and chooses which consensus solver, either the exact one for small instances or the genetic algorithm for larger problems.}
-\label{fig1}
-\end{center}
-\end{figure}
-
-\subsection*{Base Pair Tree Representation of Secondary Structure}
-% base pair tree representation vs abstract shapes
-% upsides and downsides of the
-% representation compared to abstract shape
-RNA secondary structure can represented by base pair trees, where only base pairs are considered.
-This representation is very similar to level 5 RNA abstract shape \cite{abstract_shapes}, albeit the stem length information is conserved.
-Although this representation greatly increases the size of the auxiliary space (codomain) compared to level 5 abstract shapes, it does not suffer from major changes in representation if small stem-loops are added or removed.
-
-\subsection*{Unit Tree Indel Distance}
-
-% relation to the tree edit distance (special scoring scheme)
-A tree edit distance function is used to compare base pair trees.
-
-Insertions and deletions of base pairs have unit cost and since all nodes have the same label there is no relabelling.
-
-This distance function is called the unit tree indel distance It is referred to as a tree indel distance and 
-The unit tree indel distance is a simple and intuitive distance function based on the well known tree edit distance \cite{tai, zhang_shasha}, where insertions and deletions have a unit cost.
-Since it is applied to base pair trees which are unlabelled, node relabelling comes at no cost.
-
-% interpretation of a score
-\noindent Under this scoring scheme, a distance of $n$ means that the smallest number of base pair breakage or formation to transition from a base pair tree to another is $n$.
-The distance function retains the metric properties.
-
-\noindent Overall, the unit tree indel distance is an interesting distance function with both intuitive interpretation and very useful mathematical properties.
-
-\subsection*{Minimizing Sum of Pairwise Distances}
+\noindent The consensus problem can be transformed into a MIN-SUM \textit{compact location problem} \cite{compact_location}.
+To reduce it, we add a large positive penalty $M$ to the distance of objects belonging to the same set.
+This way, any consensus using more than one object belonging to the same set is noncompetitive.
+Since the MIN-SUM compact location problem is NP-HARD, even for metric distance functions \cite{compact_location}, the use of heuristics is justified.
 
@@
 % introduce compact location problem
-
 % basic properties
@@
 
-Minimizing the sum of pairwise distance between sets of objects has been extensively investigated in the field of location theory \cite{compact_location}. Indeed, each step of the \newline
+%Minimizing the sum of pairwise distance between sets of objects has been extensively investigated in the field of location theory \cite{compact_location}. Indeed, each step of the \newline
+%
+%
+%% mapping of the consensuss problem to the compact location problem by
+%% tweaking the distance function using infinity as intra group distance
+%% may need more details to proove it, let's see what Francois thinks
+%\noindent The consensus problem described here can be reduced to an instance of the minimum average distance placement (MIN-SUM) by modifying the distance function in use such that elements belonging to the same group have infinite distance, thereby guaranteeing that a consensus will never contain more than one element of each group.
+This convenient reduction allows us to benefit from previously derived hardness results and justify the use of meta-heuristics.
+%
+%\noindent Using a metric distance function, one might think that it would be possible to reduce efficiently the search space, but even finding a near-optimal placement within a certain factor of the optimal has been proved to be NP-Hard \cite{compact_location}.
 
-% mapping of the consensuss problem to the compact location problem by
-% tweaking the distance function using infinity as intra group distance
-% may need more details to proove it, let's see what Francois thinks
-\noindent The consensus problem described here can be reduced to an instance of the minimum average distance placement (MIN-SUM) by modifying the distance function in use such that elements belonging to the same group have infinite distance, thereby guaranteeing that a consensus will never contain more than one element of each group.
-This convenient reduction allows us to benefit from previously derived hardness results and justify the use of meta-heuristics.
+\subsection*{RNA Secondary Structure and the Consensus Problem}
 
-\noindent Using a metric distance function, one might think that it would be possible to reduce efficiently the search space, but even finding a near-optimal placement within a certain factor of the optimal has been proved to be NP-Hard \cite{compact_location}.
+Most RNA secondary structure prediction tools allow for the exploration of suboptimal secondary structures.
+Very often, the experimentally determined structures of reference structures can be found close to the best structure found.
+
+When working with functionally related RNAs such as Iron Response Elements (IREs) or transfer RNAs (tRNAs), one might expect that the function can be explained by some, potentially many, shared structures.
+To this end, our goal was to devise a method capable of finding similar structures
+%citer l'article de mc-fold et Paul
+
+MC-Cons is a two-step optimization procedure (Fig.~\ref{fig1}).
+The first step is to create consensus of base pair trees using a tree edit distance function to compare them.
+This step insures that the overall topology of stems in the structures of the consensus is as similar as possible.
+The second step refines the consensus previously created by minimizing the string edit distance on the Vienna dot-bracket representation of RNA secondary structure.
+This step insures that unpaired regions match as well as possible.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[scale=0.4]{figs/mccons_flowchart2.eps}
+\caption{{\bf Flowchart illustration of MC-Cons}
+The two-step procedure is simple.
+Its input is a marna-like formatted file of RNA sequences and secondary structures in Vienna dot-bracket format.
+Its output is a marna-like formatted  The two main optimization steps are highlighted in grey.}
+\label{fig1}
+\end{center}
+\end{figure}
+
+\subsection*{Base Pair Trees and Unit Tree Indel Distance}
+% base pair tree representation vs abstract shapes
+% upsides and downsides of the
+% representation compared to abstract shape
+We define base pair trees as ordered rooted trees built from the base pairs of a RNA secondary structure.
+This representation is similar to the RNA abstract shapes level 5 \cite{abstract_shapes} with the added benefit that it preserve stems length information.
+
+% relation to the tree edit distance (special scoring scheme)
+\noindent To compare such trees, we use a standard tree edit distance \cite{zhang_shasha} along with a custom scoring scheme to compare base pair trees.
+Under this scoring scheme, insertions and deletions have unit cost.
+
+% interpretation of a score
+\noindent This distance function is intuitive.
+A distance of $n$ between two base pair trees means that it is possible to convert each tree into the other with a minimum of $n$ modifications (insertions and deletions of base pairs).
+
+%Notice that the distance function retains the metric properties (non-negativity, symmetry, subadditivity).
 
@@
 % complexity justifies meta-heuristic
-Given the complexity of the consensus problem, the use of heuristics is well justified.
+Given the complexity of the consensus problem, the use of heuristics is well justified. We still give an exact solver.
 
@@
 \section*{Results}
+
+To demonstrate the performance of MC-Cons, 
 Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
@@
 
-\section*{Software Availability} %taken from dcGOR: Software for Ontologies and Domain Annotation
+\section*{Availability and Future Directions} %taken from dcGOR: Software for Ontologies and Domain Annotation
 As open-source software, MC-Cons is freely available under the GNU LGPL v3.0 license.
-It is distributed at GitHub, https://github.com/major-lab/MC-Cons along with documentation and working examples.
-The algorithms are implemented in portable ISO 1998 C++ without external dependencies and offers a very simple command line interface.
+It is available at https://github.com/major-lab/MC-Cons along with documentation and working examples.
+The algorithms are implemented in ISO 1998 C++ without external dependencies and a basic command line interface.
 

### 2015-08-13 include/UPGMA.h
minor modifs to shape explorer and starting the UPGMA output clustering methods

@@ -1,0 +1,8 @@
+#ifndef UPGMA_H
+#define UPGMA_H
+
+
+
+
+
+#endif  // UPGMA_H

### 2015-08-13 src/UPGMA.cpp

@@ -1,0 +1,1 @@
+#include "../include/UPGMA.h"

### 2015-08-13 src/shapeExplorer.cpp

@@ -106,7 +106,7 @@ public:
         buffer << "> " << index_ << " " << tree_dist_ << " " << string_dist_ << '\n';
         for (size_t i = 0; i != structures_.size(); ++i)
         {
-            buffer << structures_[i] << " " << shapes_[i] << '\n';
+            buffer << structures_[i] << "\t" << shapes_[i] << '\n';
         }
         buffer << '\n';
         return buffer.str();

### 2015-08-13 doc/article/article.tex
did some work on the article and generated output for ires

@@
 
-We consider the problem of selecting 
-RNA secondary structural assignment is an interesting problem. Bla bla bla.
-
 % Please keep the Author Summary between 150 and 200 words
@@
 
-\subsection*{RNA Secondary Structure and the Consensus Problem}
+\subsection*{Consensus Problem and RNA Secondary Structure}
 
 Most RNA secondary structure prediction tools allow for the exploration of suboptimal secondary structures.
-Very often, the experimentally determined structures of reference structures can be found close to the best structure found.
+This is important because RNA achieve function only with multiple alternating structures \cite{need_subopts}.
+Also, the experimentally determined structures of reference structures are usually close to the best structure found.
 
-When working with functionally related RNAs such as Iron Response Elements (IREs) or transfer RNAs (tRNAs), one might expect that the function can be explained by some, potentially many, shared structures.
-To this end, our goal was to devise a method capable of finding similar structures
-%citer l'article de mc-fold et Paul
+\noindent Usually, we expect that the structures adopted by families of RNAs with similar functions such as Iron Response Elements (IREs) or transfer RNAs (tRNAs) are similar.
+To this end, our goal was to devise a method capable of finding similar structures within separate sets of suboptimal structures.
 
-MC-Cons is a two-step optimization procedure (Fig.~\ref{fig1}).
-The first step is to create consensus of base pair trees using a tree edit distance function to compare them.
+\noindent MC-Cons is a two-step optimization procedure solving two consensus problems (Fig.~\ref{fig1}).
+The first step creates consensus of base pair trees using a tree edit distance function to compare them.
 This step insures that the overall topology of stems in the structures of the consensus is as similar as possible.
@@
 Its input is a marna-like formatted file of RNA sequences and secondary structures in Vienna dot-bracket format.
-Its output is a marna-like formatted  The two main optimization steps are highlighted in grey.}
+The two main optimization steps are highlighted in grey.}
 \label{fig1}
@@
 % representation compared to abstract shape
-We define base pair trees as ordered rooted trees built from the base pairs of a RNA secondary structure.
+Base pair trees are ordered rooted trees built from the base pairs of a RNA secondary structure.
 This representation is similar to the RNA abstract shapes level 5 \cite{abstract_shapes} with the added benefit that it preserve stems length information.
@@
 A distance of $n$ between two base pair trees means that it is possible to convert each tree into the other with a minimum of $n$ modifications (insertions and deletions of base pairs).
-
-%Notice that the distance function retains the metric properties (non-negativity, symmetry, subadditivity).
+The unit tree indel distance function is metric.
 
@@
 % complexity justifies meta-heuristic
-Given the complexity of the consensus problem, the use of heuristics is well justified. We still give an exact solver.
+Two consensus solvers are available.
+Each one has the option of considering suboptimal solutions during the first optimization step.
+The first is an exact solver using a branch and bound design.
+The algorithm builds solutions incrementally and evaluates the lower bound of an incomplete solution by the summation of pairwise distances between the chosen objects and the best distance they have to remaining sets of objects to choose from.
+This algorithm works on small instances.
 
-% check which details should be given, keep it as simple as possible
-% article using metaheuristics? stuff from class maybe?
-
-% mention the interface, the solvers are given a precomputed distance matrix!
-\noindent Many simple heuristics were found to perform well within reasonable time. These algorithms all take precomputed distance matrices whose calculation is embarrassingly parallel.
-
-% for the article, we'll ignore the steepest descent solver for now, it is useless
-% version 1: Exact Version (branch & bound)
-\paragraph{Exact Version: Branch \& Bound}
-First, an exact version of the consensus solver based on a branch \& bound design was created to compare the performance of different heuristics on small instances.
-
-\noindent The lower bound of an incomplete solution is estimated by the summation of pairwise distances between the chosen objects and the best distance they have to remaining sets of objects to be chosen.
-This bound is generally quite good on small instances, but its performance degrades rapidly as the number of molecules grows.
-
-% version 2: Genetic Algorithm
-\paragraph{Heuristic: Genetic Algorithm}
-The genetic algorithm solver uses generic mutation (random substitution) and crossover operators (uniform) along with elitist selection.
-To increase convergence rate, an improvement operator using a steepest descent procedure is used.
-
-\noindent The local search is done by finding the substitution which improves the most the current sum of pairs distance.
-The operator applies a small number of iterations of local search with a certain probability.
+\noindent The second solver is an hybrid genetic algorithm which works well on medium to large instances.
+It uses generic mutation (random substitution) and crossover operators (uniform) along with an elitist selection.
+The algorithm also uses a local search procedure to increase the convergence rate.
+It finds the substitution which improves the most the current sum of pairwise distances.
+With a low probability, the improvement operator applies a small number of iterations of the local search.
 
@@
 
-To demonstrate the performance of MC-Cons, 
+\paragraph{IRES}
+
 Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
@@
 
+\bibitem{need_subopts}
+Dethoff, Elizabeth A. and Petzold, Katja and Chugh, Jeetender and Casiano-Negroni, Anette and Al-Hashimi, Hashim M. Visualizing transient low-populated structures of RNA. Nature 491, 2012
+
 \bibitem{compact_location}

### 2015-08-14 tmp/display_consensus.py
added display code in python

@@ -1,0 +1,7 @@
+"""pretty print a consensus using a minimal spanning tree"""
+
+import argparse
+from util import read_consensus_file
+import minimal_spanning_tree
+
+

### 2015-08-14 tmp/util.py

@@ -1,0 +1,478 @@
+"""Objects and functions used in multiple scripts"""
+
+
+import collections
+
+
+def is_valid_dot_bracket(dot_bracket):
+    """tests Vienna dot-bracket for illegal structure (or symbol)"""
+    counter = 0
+    for i in dot_bracket:
+        if i == '(':
+            counter += 1
+        elif i == ')':
+            counter -= 1
+        elif i != '.':  # illegal symbol
+            return False
+        if counter < 0:  # unbalanced structure
+            return False
+    if counter != 0:
+        return False  # unbalanced structure
+    return True
+
+
+def dot_bracket_to_abstract_shape(structure):
+    """Converts a Vienna dot-bracket structure into
+       its corresponding abstract shape as defined in RNAshapes.
+       Written by Stephen Leong Koan, IRIC, 2013"""
+    # the 3 levels we use are the following
+    # 1: Most accurate - all loops and all unpaired
+    # 3: Nesting pattern for all loop types but no unpaired regions
+    # 5: Most abstract - helix nesting pattern and no unpaired regions
+    list_opener = []
+
+    list_stems = []
+    i = 0
+    list_stem_start = []
+
+    # separate into stems
+    while i < len(structure):
+        if structure[i] == "(":
+            list_opener.append(i)
+            if not list_stem_start:
+                list_stem_start.append(i)
+        elif structure[i] == ")":
+            current_stem = dict(opener=[], closer=[], pairs=[])
+            while i < len(structure):
+                if structure[i] == ")":
+                    closer = list_opener.pop(-1)
+                    current_stem["opener"].append(closer)
+                    current_stem["closer"].append(i)
+                    current_stem["pairs"].append((closer, i))
+                    if closer in list_stem_start:
+                        break
+                elif structure[i] == "(":
+                    list_stem_start.append(i)
+                    i -= 1
+                    break
+                i += 1
+            list_stems.append(current_stem)
+        i += 1
+
+    # build the level 1 for each stems
+    range_occupied = []
+    dict_lvl1 = dict()
+    for stem in list_stems:
+        range_open = range(min(stem['opener']), max(stem['opener'])+1)
+        range_close = range(min(stem['closer']), max(stem['closer'])+1)
+        range_occupied.extend(range_open + range_close)
+
+        temp_lvl1_open = " "
[…]

### 2015-08-14 tmp/display_consensus.py
pushed the wrong file ...

@@ -1,7 +1,0 @@
-"""pretty print a consensus using a minimal spanning tree"""
-
-import argparse
-from util import read_consensus_file
-import minimal_spanning_tree
-
-

### 2015-08-14 src/mcconsHeuristic.cpp
reworked the command line options and added seed verification and input for the heursitic

@@ -13,18 +13,75 @@ int main(int argc, char *argv[])
 
     // create the command line parser
     OptionParser parser = OptionParser().description("MC-Cons Consensus Optimizer Using a Genetic Algorithm");
-    parser.add_option("-f", "--data").dest("dataFile").help("path to MARNA-like input file");
+
+    // I/O settings
+    parser.add_option("-i", "--data").dest("dataFile").help("path to MARNA-like input file");
+    parser.add_option("-s", "--silent").action("store_true").dest("silent").help("don't display status to stderr");
+
+    // heuristics parameters
     parser.add_option("-p", "--popsize").dest("popSize").help("genetic algorithm population size").type("size_t");
     parser.add_option("-n", "--numgen").dest("numGenerations").type("size_t").help("genetic algorithm number of generations");
-    parser.add_option("-s", "--silent").action("store_true").dest("silent").help("don't display status to stderr");
-    parser.add_option("-t", "--threshold").dest("threshold").type("size_t").help("consider up to t additional score for trees");
+
+    // threshold on tree consensus
+    parser.add_option("-t", "--threshold").dest("threshold").type("size_t").help("consider up to t additional score for tree consensus (vs optiomal one)");
+
+    // prng seeds
+    parser.add_option("--seed0").dest("seed0").type("long").help("seed 0 used by the prng");
+    parser.add_option("--seed1").dest("seed1").type("long").help("seed 1 used by the prng");
+    parser.add_option("--seed2").dest("seed2").type("long").help("seed 2 used by the prng");
+    parser.add_option("--seed3").dest("seed3").type("long").help("seed 3 used by the prng");
+    parser.add_option("--seed4").dest("seed4").type("long").help("seed 4 used by the prng");
+    parser.add_option("--seed5").dest("seed5").type("long").help("seed 5 used by the prng");
+
+
+//     parser.add_option("--sedd2").dest("seed2").type("unsigned)
     optparse::Values options = parser.parse_args(argc, argv);
 
 
     // GLOBAL SETTINGS (MODIFY AT YOUR OWN RISKS (WHICH ARE MINIMAL))
     unsigned long SEEDS[6] = {42, 42, 42, 42, 42, 42};
-//     assert((SEEDS != {0, 0, 0, 0, 0, 0}) &&
[…]

### 2015-08-16 doc/article/article.tex
removed reduction to compact location problem, was wrong

@@
 
-Given $n$ sets of $ m_1, m_2... m_n$ objects, the consensus problem is to find a set of $n$ objects, one per set, minimizing the sum of their pairwise distance.
+Given $k$ sets of $ m_1, m_2...
+m_n$ objects, the consensus problem is to find a set of $k$ objects, one per set, minimizing the sum of their pairwise distance.
+This is modelled as finding the smallest cycle in a complete $k$-partite graph.
+
+% http://stackoverflow.com/questions/17168049/max-weight-k-clique-in-a-complete-k-partite-graph
+% http://cstheory.stackexchange.com/questions/30679/k-clique-in-k-partite-graph
+% http://cstheory.stackexchange.com/questions/30605/many-one-reduction-from-inequality-problem-to-equality-problem
+\noindent The consensus problem is reformulated as the problem of finding a minimal $k$-clique in a complete weighted $k$-partite graph where $k$ is a free parameter.
+This problem is known to be intractable
 

### 2015-08-27 doc/article/article.tex
worked a bit on the wording and minor improvements to mccons

@@
 \subsection*{Consensus Problem}
-
-Given $k$ sets of $ m_1, m_2...
-m_n$ objects, the consensus problem is to find a set of $k$ objects, one per set, minimizing the sum of their pairwise distance.
-This is modelled as finding the smallest cycle in a complete $k$-partite graph.
-
 % http://stackoverflow.com/questions/17168049/max-weight-k-clique-in-a-complete-k-partite-graph
@@
 % http://cstheory.stackexchange.com/questions/30605/many-one-reduction-from-inequality-problem-to-equality-problem
-\noindent The consensus problem is reformulated as the problem of finding a minimal $k$-clique in a complete weighted $k$-partite graph where $k$ is a free parameter.
-This problem is known to be intractable
+Given $k$ sets of $m_0, m_1... m_k$ objects, the consensus problem is to find a set of $k$ objects, one per set, minimizing the sum of their pairwise distance.
+This is equivalent to the maximum edge-weighted clique problem in a complete $k$-partite graph. This problem is NP-hard and this justifies the use of heuristic methods on larger instances.
 
-\noindent The consensus problem can be transformed into a MIN-SUM \textit{compact location problem} \cite{compact_location}.
+%\noindent The consensus problem can be transformed into a MIN-SUM \textit{compact location problem} \cite{compact_location}.
 To reduce it, we add a large positive penalty $M$ to the distance of objects belonging to the same set.
@@
 Most RNA secondary structure prediction tools allow for the exploration of suboptimal secondary structures.
-This is important because RNA achieve function only with multiple alternating structures \cite{need_subopts}.
-Also, the experimentally determined structures of reference structures are usually close to the best structure found.
+This is important because some RNA achieve function only with multiple structures \cite{need_subopts}.
+Also, the experimentally determined structures of reference structures are often close to the best structure predicted.
 

### 2015-08-27 src/ConsensusClusterMap.py

@@ -13,10 +13,8 @@ from Util import *
 seaborn.set(font="monospace")
 
 
-
-
 class CustomClusterGrid(seaborn.matrix.Grid):
-    def __init__(self, data, pivot_kws=None, figsize=None, row_colors=None, col_colors=None):
+    def __init__(self, data, pivot_kws=None, figsize=(10, 10), row_colors=None, col_colors=None):
         """Grid object for organizing clustered heatmap input on to axes"""
 
         if isinstance(data, pd.DataFrame):
@@ -24,16 +22,8 @@ class CustomClusterGrid(seaborn.matrix.Grid):
         else:
             self.data2d = pd.DataFrame(data)
 
-        if figsize is None:
-            width, height = 10, 10
-            figsize = (width, height)
         self.fig = plt.figure(figsize=figsize)
-
-        if row_colors is not None:
-            row_colors = _convert_colors(row_colors)
         self.row_colors = row_colors
-        if col_colors is not None:
-            col_colors = _convert_colors(col_colors)
         self.col_colors = col_colors
 
         width_ratios = self.dim_ratios(self.row_colors,
@@ -46,7 +36,8 @@ class CustomClusterGrid(seaborn.matrix.Grid):
         nrows = 3 if self.col_colors is None else 4
         ncols = 3 if self.row_colors is None else 4
 
-        self.gs = matplotlib.gridspec.GridSpec(nrows, ncols, wspace=0.01, hspace=0.01,
+        self.gs = matplotlib.gridspec.GridSpec(nrows, ncols, wspace=0.01,
+                                               hspace=0.01,
                                                width_ratios=width_ratios,
                                                height_ratios=height_ratios)
 
@@ -65,14 +56,16 @@ class CustomClusterGrid(seaborn.matrix.Grid):
             self.ax_col_colors = self.fig.add_subplot(
                 self.gs[nrows - 2, ncols - 1])
 
-        self.ax_heatmap = self.fig.add_subplot(self.gs[nrows - 1, ncols - 1])
+        self.ax_heatmap = self.fig.add_subplot(self.gs[nrows - 1,
+                                                       ncols - 1])
 
 
         self.dendrogram_row = None
         self.dendrogram_col = None
 
 
-    def dim_ratios(self, side_colors, axis, figsize, side_colors_ratio=0.05):
+    def dim_ratios(self, side_colors, axis,
+                   figsize, side_colors_ratio=0.05):
         figdim = figsize[axis]
[…]

### 2015-08-27 src/MCCONS.cpp

@@ -116,23 +116,6 @@ void MCCONS(std::string path,
     }
     tree_consensus = tree_consensus_tmp;
 
-    // PRINT THE TREE CONSENSUS
-    std::cerr << std::endl << "Tree Consensus Solutions" << std::endl;
-    if (! silent_output)
-    {
-        int index;
-        for(size_t i = 0; i != tree_consensus.size(); ++i)
-        {
-            std::cerr << "> " << i << " " << tree_consensus[i].get_score()/num_comparisons << std::endl;
-            for(size_t j = 0; j != tree_consensus[i].get_genes().size(); ++j)
-            {
-                index = tree_consensus[i].get_genes()[j];
-                std::cerr << tree_problem.get_objects()[index].get_brackets() << std::endl;
-            }
-            std::cerr << std::endl;
-        }
-    }
-
 
     // PHASE 1.5: FILTERING
     std::vector<std::vector< std::vector<std::string> > > prob2_data = std::vector< std::vector< std::vector<std::string> > >();

### 2015-08-27 src/SolverHeuristic.cpp

@@ -5,7 +5,7 @@ using std::vector;
 
 
 // constructors and destructors
-SolverHeuristic::SolverHeuristic(// misc parameters
+SolverHeuristic::SolverHeuristic(
                    bool silent,
 
                    // GA settings
@@ -43,18 +43,14 @@ Solution uniform_crossover(Solution parent1,
                            RngStream* prng)
 {
     // uniform crossover operator over the genes of two solutions
-    vector<int> inherited_genes = vector<int>();
-    double prob; // Uniform(0, 1)
-    for (size_t index = 0; index != parent1.get_genes().size(); ++index)
+    vector<int> inherited_genes (parent1.get_genes());
+    size_t gene_length = parent1.get_genes().size();
+
+    for (size_t index = 0; index != gene_length; ++index)
     {
-        prob = prng->RandU01();
-        if (prob < crossover_probability)
+        if (prng->RandU01() < crossover_probability)
         {
-            inherited_genes.push_back(parent1.get_genes()[index]);
-        }
-        else
-        {
-            inherited_genes.push_back(parent2.get_genes()[index]);
+            inherited_genes[index] = parent2.get_genes()[index];
         }
     }
     return Solution(inherited_genes);
@@ -62,27 +58,20 @@ Solution uniform_crossover(Solution parent1,
 
 
 Solution uniform_mutate(Solution& sol,
-                         std::vector<Range> ranges,
-                         RngStream *prng,
-                         double mutation_probability=0.1)
+                        std::vector<Range> ranges,
+                        double mutation_probability,
+                        RngStream *prng)
 {
     // mutate the solution by simply swapping with a probability
-    double prob;
-    vector<int> mutated_genes = vector<int>(sol.get_genes().size(), 0);
-
-    for (size_t index = 0; index != sol.get_genes().size(); ++index)
+    vector<int> mutated_genes (sol.get_genes());
+    size_t gene_size = mutated_genes.size();
+    for (size_t index = 0; index != gene_size; ++index)
     {
-        prob = prng->RandU01();
-        if (prob < mutation_probability)
+        if (prng->RandU01() < mutation_probability)
         {
             // exchange for a random gene within the same range
             mutated_genes[index] = prng->RandInt(ranges[index].low,
                                                  ranges[index].high-1);
-        }
-        else
-        {
-            // keep the same gene
-            mutated_genes[index] = sol.get_genes()[index];
         }
     }
     return Solution(mutated_genes);
[…]

### 2015-09-05 doc/article/article.tex
saving for other computer

@@
 
-\noindent Usually, we expect that the structures adopted by families of RNAs with similar functions such as Iron Response Elements (IREs) or transfer RNAs (tRNAs) are similar.
+\noindent Usually, we expect that the structures adopted by families of RNAs with similar functions such as Iron Response Elements (IREs) or transfer RNAs (tRNAs) to be similar.
 To this end, our goal was to devise a method capable of finding similar structures within separate sets of suboptimal structures.

### 2015-09-05 src/Util.py

@@ -448,7 +448,7 @@ def read_consensus_file(file_name,
                 consensus.append(Consensus(number,
                                           subopts))
     else:
-        current_index = -2
+        current_index = -1
         with open(file_name) as data_file:
             for line in data_file:
                 # starting new consensus
@@ -470,8 +470,7 @@ def read_consensus_file(file_name,
 
             # last consensus (if the file wasn't empty)
             if subopts:
-                consensus.append(Consensus(number,
-                                          subopts))
+                return Consensus(number, subopts)
     return consensus
 
 

### 2015-10-07 tmp/ConsensusClusterMap.py
updated presentation version 1, still needs feedback

@@ -1,0 +1,181 @@
+#! /usr/bin/env python
+"""save a consensus into heatmap + dendrogram form (with seaborn)"""
+
+import argparse
+import matplotlib
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import scipy.cluster
+import seaborn as sns
+from Util import *
+
+sns.set(font="monospace")
+sns.set_style('white')
+
+
+def compute_weighted_distances(consensus,
+                               shape_weight,
+                               tree_weight,
+                               string_weight):
+    """converts the multiple distances used into a single distance
+       matrix by a simple weighted sum (many possibilities)"""
+
+    # first, we'd like to establish a distance between shapes
+    num_elems = len(consensus.subopts)
+    shape_dists = np.zeros((num_elems, num_elems))
+
+    # calculate the unit tree edit distance on the shapes
+    for index_1 in range(num_elems - 1):
+        (level_5_1, _, _) = consensus.abstract_shapes[index_1]
+        dot_bracket_1 = level_5_1.replace("[", "(")
+        dot_bracket_1 = dot_bracket_1.replace("]", ")")
+        tree_1 = dot_bracket_to_tree(dot_bracket_1)
+        for index_2 in range(index_1 + 1, num_elems):
+            (level_5_2, _, _) = consensus.abstract_shapes[index_2]
+            dot_bracket_2 = level_5_2.replace("[", "(")
+            dot_bracket_2 = dot_bracket_2.replace("]", ")")
+            tree_2 = dot_bracket_to_tree(dot_bracket_2)
+            # assign it to the matrix
+            dist = unlabeled_distance(tree_1, tree_2)
+            shape_dists[index_1][index_2] = dist
+            shape_dists[index_2][index_1] = dist
+
+    # add the 3 matrices with proper weights
+    new_matrix = np.zeros((num_elems, num_elems))
+    for index_1 in range(num_elems):
+        for index_2 in range(num_elems):
+            shape_score = shape_weight * shape_dists[index_1][index_2]
+            tree_score = tree_weight * consensus.tree_dists[index_1][index_2]
+            string_score = (string_weight *
+                            consensus.string_dists[index_1][index_2])
+            total_score = tree_score + shape_score + string_score
+            new_matrix[index_1][index_2] = total_score
+    return new_matrix
+
+
+def get_dendrogram(distance_matrix, leaf_label_func, cluster_method="average"):
+    """returns the order in which the consensus should be presented"""
+    # perform the hierarchical clustering
[…]

### 2015-10-08 include/ConsensusProblem.h
removed include files (useless now)

@@ -1,43 +1,0 @@
-#ifndef CONSENSUSPROBLEM_H
-#define CONSENSUSPROBLEM_H
-
-#include <algorithm>
-#include <limits>
-#include <map>
-#include <string>
-#include <vector>
-
-#include "Tree.h"
-
-
-struct Range
-{  // to keep the boundaries of each group of objects
-    int low;
-    int high;
-};
-
-
-template <class T>
-class ConsensusProblem
-{  // abstract consensus problem class
-public:
-    // constructor and destructor
-    ConsensusProblem(std::vector< std::vector<T> > data,
-                     int (*distance_function)(T a, T b));
-    ~ConsensusProblem(){ }
-
-    // getters and setters
-    const std::vector<T> & get_objects();
-    const std::vector<Range> & get_ranges();
-    const std::vector< std::vector<double> > & get_distance_matrix();
-
-
-private:
-    std::vector<T> objects_;
-    std::vector<Range> ranges_;
-    std::vector< std::vector<double> > distance_matrix_;
-};
-
-
-#endif // CONSENSUSPROBLEM_H
-

### 2015-10-08 include/Distances.h

@@ -1,26 +1,0 @@
-#ifndef DISTANCES_H
-#define DISTANCES_H
-
-#include <string>
-#include <vector>
-
-#include "RNA2D.h"
-#include "Tree.h"
-
-
-// unit tree indel distance on strings (dot brackets)
-int unit_tree_indel_distance_strings(const std::string string1,
-                                     const std::string string2);
-
-
-// unit tree indel distance on trees
-int unit_tree_indel_distance_trees(const OrderedLabeledTree tree1,
-                                   const OrderedLabeledTree tree2);
-
-
-// string edit distance on ... strings
-int string_edit_distance(std::string string1,
-                         std::string string2);
-
-#endif
-

### 2015-10-08 include/MCCONS.h

@@ -1,21 +1,0 @@
-#ifndef MCCONS_H
-#define MCCONS_H
-
-#include <utility>
-#include <algorithm>
-#include <string>
-
-
-#include "Distances.h"
-#include "ProgressBar.h"
-#include "Readers.h"
-#include "RNA2D.h"
-#include "Solver.h"
-
-
-// The whole two-step optimization procedure is done by this
-// function which is controlled by feeding different solvers
-void MCCONS(std::string path, Solver* tree_solver, Solver* dotbracket_solver);
-
-
-#endif  // MCCONS_H

### 2015-10-08 include/ProgressBar.h

@@ -1,28 +1,0 @@
-#ifndef PROGRESSBAR_H
-#define PROGRESSBAR_H
-
-#include <string>
-#include <iostream>
-
-#include <assert.h>
-
-class ProgressBar
-{  // simple progress bar
-
-public:
-    // constructors and destructor
-    ProgressBar(std::string process_name="", size_t bar_length=40);
-    ~ProgressBar();
-
-    // updates the progress displayed
-    void update(float percentage);
-
-    // displays 100% status, should be called immediately after last iteration
-    void clean();
-
-private:
-    size_t bar_length_;
-    std::string process_name_;
-};
-
-#endif // PROGRESSBAR_H

### 2015-10-08 include/RNA2D.h

@@ -1,18 +1,0 @@
-#ifndef RNA2D_H
-#define RNA2D_H
-
-#include <algorithm>
-#include <assert.h>
-#include <string>
-#include <vector>
-
-// asserts that the Vienna dot bracket is valid
-// (balanced, no illegal symbols)
-bool is_valid_dot_bracket(std::string dot_bracket);
-
-
-// removes the dots from the vienna dot bracket
-std::string only_paired(std::string dot_bracket);
-
-
-#endif  // RNA2D_H

### 2015-10-08 include/RNAshapes.h

@@ -1,18 +1,0 @@
-
-
-#ifndef RNASHAPES_H
-#define RNASHAPES_H
-
-#include <string>
-#include <vector>
-#include <algorithm>
-#include <map>
-#include <deque>
-#include <cassert>
-
-
-// only level 1, 3 and 5 implemented
-std::string RNAshapes(std::string dot_bracket, int level);
-
-
-#endif // RNASHAPES_H

### 2015-10-08 include/Readers.h

@@ -1,22 +1,0 @@
-#ifndef READERS_H
-#define READERS_H
-
-
-#include <stdlib.h> // for exit
-
-
-#include <fstream>  // ifstream
-#include <iostream> // file reading
-#include <string>
-#include <vector>
-
-
-// read MC-Cons input files (marna-like formatted lists of Vienna dot brackets)
-std::vector<std::vector<std::string> > read_marna_file(std::string file_name);
-
-
-// read MC-Cons output files (consensus files)
-std::vector<std::vector<std::string> > read_consensus_file(std::string file_name);
-
-
-#endif // READERS_H

### 2015-10-08 include/RngStream.h

@@ -1,78 +1,0 @@
-
-#ifndef RNGSTREAM_H
-#define RNGSTREAM_H
-
-#include <string>
-
-class RngStream
-{
-public:
-
-    RngStream (const char *name = "");
-
-
-    static bool SetPackageSeed (const unsigned long seed[6]);
-
-
-    void ResetStartStream ();
-
-
-    void ResetStartSubstream ();
-
-
-    void ResetNextSubstream ();
-
-
-    void SetAntithetic (bool a);
-
-
-    void IncreasedPrecis (bool incp);
-
-
-    bool SetSeed (const unsigned long seed[6]);
-
-
-    void AdvanceState (long e, long c);
-
-
-    void GetState (unsigned long seed[6]) const;
-
-
-    void WriteState () const;
-
-
-    void WriteStateFull () const;
-
-
-    double RandU01 ();
-
-
-    int RandInt (int i, int j);
-
-
-
-private:
-
-    double Cg[6], Bg[6], Ig[6];
-
-
-    bool anti, incPrec;
-
-
-    std::string name;
-
-
-    static double nextSeed[6];
-
-
-    double U01 ();
-
-
-    double U01d ();
-
-
-};
-
-#endif
-
-

### 2015-10-08 include/ShapeExplorer.h

@@ -1,21 +1,0 @@
-#ifndef SHAPEEXPLORER_H
-#define SHAPEEXPLORER_H
-
-
-#include <assert.h>
-
-
-#include <string>
-#include <vector>
-#include <map>
-#include <iostream>
-#include <sstream>
-
-
-#include "Distances.h"
-#include "OptionParser.h"
-#include "RNAshapes.h"
-#include "Readers.h"
-
-
-#endif // SHAPEEXPLORER_H

### 2015-10-08 include/Solution.h

@@ -1,36 +1,0 @@
-#ifndef SOLUTION_H
-#define SOLUTION_H
-
-#include <iostream>  // <<
-#include <vector>
-#include <limits>    // infinity for default score
-
-class Solution
-{
-public:
-    // constructors and destructors
-    Solution(std::vector<int> genes=std::vector<int>(), double score=std::numeric_limits<double>::infinity());
-    Solution(const Solution& other);
-    ~Solution();
-
-    // getters and setters
-    const std::vector<int> & get_genes() const;
-    void set_gene(int position, int new_gene);
-
-    double get_score() const;
-    void set_score(double new_score);
-
-    // some comparison operators
-    bool operator <(const Solution &other) const;
-    bool operator ==(const Solution &other) const;
-
-    // pretty printer
-    friend std::ostream& operator<< (std::ostream &out, Solution &sol);
-
-private:
-    std::vector<int> genes_;
-    double score_;
-};
-
-
-#endif // SOLUTION_H

### 2015-10-08 include/Solver.h

@@ -1,67 +1,0 @@
-#ifndef SOLVER_H
-#define SOLVER_H
-
-#include <assert.h>
-
-#include <limits>
-#include <vector>
-
-#include "ConsensusProblem.h"
-#include "ProgressBar.h"
-#include "RngStream.h"
-#include "Solution.h"
-#include "Tree.h"
-
-
-
-class Solver
-{  // abstract base class
-
-public:
-    // ctor
-    Solver(bool silent) { silent_ = silent; }
-
-    // silent status
-    void flip_silent() { silent_ = !silent_; }
-    bool is_silent() { return silent_; }
-
-    // method to implement for inheriting classes
-    virtual std::vector<Solution> solve(std::vector< std::vector<double> > distance_matrix,
-                                        std::vector<Range> ranges) const = 0; // pure virtual
-
-
-protected:
-    bool silent_;
-};
-
-
-
-struct int_double
-{  // simple data holder
-    int best_gene;
-    double cost_difference;
-};
-
-
-// calculates and assigns the sum of pairwise distances
-// of a solution and assign it as score_
-void assign_pairwise_distance_score(Solution& sol, std::vector< std::vector<double> > distance_matrix);
-
-
-// randomly chooses numbers fitting the boundaries of the ranges
-std::vector<int> select_random_genes(std::vector<Range> ranges, RngStream *prng);
-
-
-// create a list of Solutions generated randomly
-std::vector<Solution> initialize_population(std::vector<Range> ranges, int pop_size, RngStream *prng);
-
-
-// applies the steepest descent procedure until either
-// a local minima is found or max_iter iterations of the substitution are done
-void steepest_improvement(Solution& sol,
-                          std::vector< std::vector<double> > distance_matrix,
-                          std::vector<Range> ranges,
-                          int max_iter=std::numeric_limits<int>::infinity());
-
-
-#endif  // SOLVER_H

### 2015-10-08 include/SolverExact.h

@@ -1,36 +1,0 @@
-#ifndef SOLVEREXACT_H_
-#define SOLVEREXACT_H_
-
-
-#include <assert.h>
-
-#include <algorithm>
-#include <limits>
-#include <queue>
-
-#include "Solution.h"
-#include "Solver.h"
-#include "SubSolution.h"
-
-
-class SolverExact: public Solver
-{  // Branch & Bound solver
-
-public:
-    // constructors and destructors
-    SolverExact(double suboptimal_threshold, bool silent);
-    ~SolverExact();
-
-
-    // solver call
-    std::vector<Solution> solve(std::vector< std::vector<double> > distance_matrix,
-                                std::vector<Range> ranges) const;
-
-
-private:
-    double suboptimal_threshold_;
-};
-
-
-#endif // SOLVEREXACT_H_
-

### 2015-10-08 include/SolverHeuristic.h

@@ -1,52 +1,0 @@
-#ifndef SOLVERHEURISTIC_H_
-#define SOLVERHEURISTIC_H_
-
-#include <algorithm>
-#include <assert.h>
-
-
-#include "Solver.h"  // abstract class and useful functions
-
-
-class SolverHeuristic : public Solver
-{  // hybrid genetic algorithm / steepest descent
-
-public:
-    // constructors and destructors
-    SolverHeuristic(// generic information
-             bool silent,
-
-             // heuristic parameters
-             int population_size,
-             int num_generations,
-             int improvement_depth,
-             int elite_size,
-             double distance_threshold,
-
-             double crossover_prob,
-             double mutation_prob,
-             double improvement_prob,
-             unsigned long seeds[6]);
-    ~SolverHeuristic();
-
-    // solver call
-    std::vector<Solution> solve(std::vector< std::vector<double> > distance_matrix,
-                                std::vector<Range> ranges) const;
-
-
-private:
-    int population_size_;
-    int elite_size_;
-    int num_generations_;
-    int improvement_depth_;
-
-    double distance_threshold_;
-    double crossover_prob_;
-    double mutation_prob_;
-    double improvement_prob_;
-    unsigned long * seeds_;
-};
-
-
-#endif // SOLVERHEURISTIC_H_
-

### 2015-10-08 include/SubSolution.h

@@ -1,60 +1,0 @@
-#ifndef SOLUTIONNODE_H
-#define SOLUTIONNODE_H
-
-
-#include <algorithm>
-#include <iostream>
-#include <vector>
-#include <limits>
-
-#include "Solution.h"
-
-
-struct RangeIndex
-{  // modified range holder
-    int index;
-    int low;
-    int high;
-};
-
-
-class SubSolution
-{  // special solution type used for Branch & Bound
-
-public:
-    // constructor and destructor
-    SubSolution(std::vector<int> genes,
-                 std::vector<RangeIndex> ranges,
-                 double score = std::numeric_limits<double>::max());
-    SubSolution(const SubSolution& other);
-    ~SubSolution();
-
-    // getters and setters
-    const std::vector<int> & get_genes() const;
-    void set_gene(int position, int new_gene);
-
-    std::vector<RangeIndex> get_ranges() const;
-
-    double get_score() const;
-    void set_score(std::vector< std::vector<double> > elem_range_dm,
-                   std::vector< std::vector<double> > distance_matrix);
-
-    // return a copy of it as a solution
-    Solution return_solution();
-
-    // some comparison operators
-    bool operator <(const SubSolution &other) const;
-    bool operator ==(const SubSolution &other);
-
-    // pretty printer
-    friend std::ostream& operator<< (std::ostream &out, const SubSolution &sol);
-
-
-private:
-    std::vector<int> genes_;
-    std::vector<RangeIndex> ranges_;
-    double score_;
-};
-
-
-#endif

### 2015-10-08 include/Tree.h

@@ -1,106 +1,0 @@
-#ifndef TREE_H
-#define TREE_H
-
-#include <assert.h>
-
-#include <algorithm>
-#include <cstddef>
-#include <iostream>
-#include <map>
-#include <string>
-#include <vector>
-
-
-// ====================================NODE====================================
-
-
-class Node
-{   // used for trees
-public:
-    // constructor
-    Node(Node* parent, char label)
-    {
-        // constructor
-        if (parent)
-        {
-            parent->add_child(this);
-        }
-        parent_ = parent;
-        label_ = label;
-        index_ = -1;
-        children_ = std::vector<Node*>();
-    };
-
-    // destructor
-    ~Node(){};
-
-    // getters
-    char get_label() const { return label_; }
-    int get_index()  const { return index_; }
-    Node* get_parent() { return parent_; }
-    const std::vector<Node*> & get_children() { return children_; }
-
-    // setters
-    void set_parent(Node* parent) { parent_ = parent; }
-    void set_label(char label) { label_ = label; }
-    void set_index(int index)  { index_ = index; }
-
-    // modifier
-    void add_child(Node* child) { children_.push_back(child); }
-
-    // helper
-    int get_leftmost_descendent_index();
-
-private:
-    Node* parent_;
-    char label_;
-    int index_;
-    std::vector<Node*> children_;
-};
-
-
-// ============================ORDERED LABELED TREE============================
-
-
-class OrderedLabeledTree
-{   // abstract base class for RNA trees
-
-public:
-    // ctor & dtor
-    OrderedLabeledTree(std::string string_repr, // representation
-                       char add_branch_symbol,  // + and down last added
-                       char unbranch_symbol,    // up
-                       char add_leaf_symbol);   // +
-    ~OrderedLabeledTree();
-
-    // accessors
-    std::vector<int> &  get_leftmost_descendents() { return leftmost_descendents_; 
-    std::vector<int> &  get_keyroots() { return keyroots_; }
-    std::vector<char> & get_labels() { return postorder_labels_; }
-    std::string get_string() { return string_repr_; }
-    char get_add_branch_symbol() { return add_branch_symbol_; }
-    char get_unbranch_symbol()  { return unbranch_symbol_; }
-    char get_add_leaf_symbol() { return add_leaf_symbol_; }
-
-        // friend functions
[…]

### 2015-10-08 python/Util.py
moved python sources to new directory

@@ -1,0 +1,477 @@
+"""Objects and functions used in multiple scripts"""
+
+
+import collections
+
+
+def is_valid_dot_bracket(dot_bracket):
+    """tests Vienna dot-bracket for illegal structure (or symbol)"""
+    counter = 0
+    for i in dot_bracket:
+        if i == '(':
+            counter += 1
+        elif i == ')':
+            counter -= 1
+        elif i != '.':  # illegal symbol
+            return False
+        if counter < 0:  # unbalanced structure
+            return False
+    if counter != 0:
+        return False  # unbalanced structure
+    return True
+
+
+def dot_bracket_to_abstract_shape(structure):
+    """Converts a Vienna dot-bracket structure into
+       its corresponding abstract shape as defined in RNAshapes.
+       Written by Stephen Leong Koan, IRIC, 2013"""
+    # the 3 levels we use are the following
+    # 1: Most accurate - all loops and all unpaired
+    # 3: Nesting pattern for all loop types but no unpaired regions
+    # 5: Most abstract - helix nesting pattern and no unpaired regions
+    list_opener = []
+
+    list_stems = []
+    i = 0
+    list_stem_start = []
+
+    # separate into stems
+    while i < len(structure):
+        if structure[i] == "(":
+            list_opener.append(i)
+            if not list_stem_start:
+                list_stem_start.append(i)
+        elif structure[i] == ")":
+            current_stem = dict(opener=[], closer=[], pairs=[])
+            while i < len(structure):
+                if structure[i] == ")":
+                    closer = list_opener.pop(-1)
+                    current_stem["opener"].append(closer)
+                    current_stem["closer"].append(i)
+                    current_stem["pairs"].append((closer, i))
+                    if closer in list_stem_start:
+                        break
+                elif structure[i] == "(":
+                    list_stem_start.append(i)
+                    i -= 1
+                    break
+                i += 1
+            list_stems.append(current_stem)
+        i += 1
+
+    # build the level 1 for each stems
+    range_occupied = []
+    dict_lvl1 = dict()
+    for stem in list_stems:
+        range_open = range(min(stem['opener']), max(stem['opener'])+1)
+        range_close = range(min(stem['closer']), max(stem['closer'])+1)
+        range_occupied.extend(range_open + range_close)
+
+        temp_lvl1_open = " "
[…]

### 2015-10-08 python/ConsensusCluster.py
upgraded presentation and modified clustering script

@@ -54,14 +54,28 @@ def compute_weighted_distances(consensus,
     return new_matrix
 
 
-def get_dendrogram(distance_matrix, leaf_label_func, cluster_method="average"):
+def get_dendrogram(Z, labels):
+    """returns the order in which the consensus should be presented"""
+    # perform the hierarchical clustering
+    return scipy.cluster.hierarchy.dendrogram(Z,  orientation="right", labels=labels, leaf_font_size=8)
+
+
+def get_order_linkage(distance_matrix, cluster_method="average"):
     """returns the order in which the consensus should be presented"""
     # perform the hierarchical clustering
     mat_ut = distance_matrix[np.triu_indices(distance_matrix.shape[0], 1)]
     Z = scipy.cluster.hierarchy.linkage(mat_ut, method=cluster_method)
 
-    return scipy.cluster.hierarchy.dendrogram(Z, leaf_label_func=leaf_label_func, orientation="left")
-
+    # convert the linkage information to a dendogram ordering
+    # http://stackoverflow.com/questions/12572436/calculate-ordering-of-dendrogram-leaves
+    n = len(Z) + 1
+    cache = dict()
+    for k in range(len(Z)):
+        c1, c2 = int(Z[k][0]), int(Z[k][1])
+        c1 = [c1] if c1 < n else cache.pop(c1)
+        c2 = [c2] if c2 < n else cache.pop(c2)
+        cache[n+k] = c1 + c2
+    return (cache[2*len(Z)], Z)
 
 #def len_digit(integer):
     #"""calculates the number of digits in an integer number"""
@@ -167,11 +181,12 @@ if __name__ == '__main__':
 
 
     FIGURE = plt.figure(figsize=(ARGS.x_size, ARGS.y_size))
-    LABELS = [CONSENSUS.subopts[k] for k in range(len(CONSENSUS.subopts))]
-    def getLab(i):
-        return LABELS[i]
-    DENDROGRAM = get_dendrogram(DISTANCE_MATRIX, leaf_label_func=getLab)
-    plt.xticks(rotation=90)
+    LABELS = [k for k in range(len(CONSENSUS.subopts))]
+    ORDER, LINKAGE = get_order_linkage(DISTANCE_MATRIX)
+    for i in ORDER:
+        print CONSENSUS.subopts[i]
+    DENDROGRAM = get_dendrogram(LINKAGE, LABELS)
+  #  plt.xticks(rotation=90)
[…]

### 2015-10-09 src/src/mccons/executables/MCConsHeuristic.java
saving presentation for remote

@@ -48,20 +48,20 @@ public final class MCConsHeuristic {
                 'p', "popSize", "Genetic algorithm population size"));
 
 
-        parser.registerParameter(new FlaggedOption("numGenerations", JSAP.INTEGER_PARSER, "500", false,
+        parser.registerParameter(new FlaggedOption("numGenerations", JSAP.INTEGER_PARSER, "1000", false,
                 'n', "numGenerations", "Number of generations to evaluate"));
 
 
-        parser.registerParameter(new FlaggedOption("eliteSize", JSAP.INTEGER_PARSER, "30", false,
+        parser.registerParameter(new FlaggedOption("eliteSize", JSAP.INTEGER_PARSER, "50", false,
                 JSAP.NO_SHORTFLAG, "eliteSize", "Size of the elite in the genetic algorithm"));
 
 
-        parser.registerParameter(new FlaggedOption("improvementProbability", JSAP.DOUBLE_PARSER, "0.05",
+        parser.registerParameter(new FlaggedOption("improvementProbability", JSAP.DOUBLE_PARSER, "0.03",
                 false, JSAP.NO_SHORTFLAG, "improvementProbability",
                 "Probability of applying steepest descent on a solution"));
 
 
-        parser.registerParameter(new FlaggedOption("improvementDepth", JSAP.INTEGER_PARSER, "2",
+        parser.registerParameter(new FlaggedOption("improvementDepth", JSAP.INTEGER_PARSER, "4",
                 false, JSAP.NO_SHORTFLAG, "improvementDepth",
                 "Number of iterations applied each time the steepest descent is called"));
 
@@ -70,7 +70,7 @@ public final class MCConsHeuristic {
                 false, JSAP.NO_SHORTFLAG, "crossoverProbability",
                 "Probability of applying uniform crossover on a child solution"));
 
-        parser.registerParameter(new FlaggedOption("crossoverMixingRatio", JSAP.DOUBLE_PARSER, "0.5",
+        parser.registerParameter(new FlaggedOption("crossoverMixingRatio", JSAP.DOUBLE_PARSER, "0.1",
                 false, JSAP.NO_SHORTFLAG, "crossoverMixingRatio",
                 "Mixing ratio used for uniform crossover between two parent solutions"));
 
@@ -79,7 +79,7 @@ public final class MCConsHeuristic {
                 false, JSAP.NO_SHORTFLAG, "mutationProbability",
[…]

### 2015-10-09 src/src/mccons/solvers/MCCons.java

@@ -4,6 +4,7 @@ import mccons.repr.ConsensusProblem;
 import mccons.repr.OrderedRootedLabeledTree;
 import mccons.repr.cost.CostFunction;
 import mccons.repr.cost.DistanceFunction;
+import mccons.repr.cost.StringEditDistance;
 import mccons.rna.RNAshapes;
 import mccons.util.Readers;
 import mccons.repr.cost.TreeEditDistance;
@@ -44,13 +45,13 @@ public class MCCons {
         public int shapeLevel;
         public double mult1, mult2, mult3;
 
-        shapeAndSkeletonDist(int shapeLevel_, double mult1_, double mult2_, double mult3_)
+        shapeAndSkeletonDist(double mult1_, double mult2_)
         {
-            assert(shapeLevel_ == 1 || shapeLevel_ == 3 || shapeLevel_ ==5);
-            shapeLevel = shapeLevel_;
+            //assert(shapeLevel_ == 1 || shapeLevel_ == 3 || shapeLevel_ ==5);
+            //shapeLevel = shapeLevel_;
             mult1 = mult1_;
             mult2 = mult2_;
-            mult3 = mult3_;
+            //mult3 = mult3_;
         }
 
         public Double get(String dotBracket1, String dotBracket2)
@@ -58,6 +59,7 @@ public class MCCons {
             OrderedRootedLabeledTree tree1 = new OrderedRootedLabeledTree(dotBracket1, '(', ')' ,'?');
             OrderedRootedLabeledTree tree2 = new OrderedRootedLabeledTree(dotBracket2, '(', ')' ,'?');
 
+            /*
             OrderedRootedLabeledTree completeTree1 = new OrderedRootedLabeledTree(dotBracket1, '(', ')', '.');
             OrderedRootedLabeledTree completeTree2 = new OrderedRootedLabeledTree(dotBracket2, '(', ')', '.');
 
@@ -66,17 +68,20 @@ public class MCCons {
 
             OrderedRootedLabeledTree shapeTree1 = new OrderedRootedLabeledTree(shape1, '[', ']', '_');
             OrderedRootedLabeledTree shapeTree2 = new OrderedRootedLabeledTree(shape2, '[', ']', '_');
-
+*/
             UnitCost indel = new UnitCost();
             SubstitutionCost sub = new SubstitutionCost();
+
             TreeEditDistance dist = new TreeEditDistance(indel, indel, sub);
+            StringEditDistance stringDist = new StringEditDistance();
 
[…]
